---
title: "Crypto_EEG_CANTAB_paper"
author: "Eily Sullivan"
date: "5/2/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages, include=FALSE}

rm(list=ls())
library(tidyverse)
library(haven)
library(dplyr)
library(foreign)
library(lmtest)
library(lme4)
library(lmerTest)
library(ggplot2)
#install.packages("standardize")
library(standardize)
library(ggthemes)
library(texreg)
library(rockchalk)
library(stargazer)
#install.packages("Hmisc")
library(Hmisc)
library(writexl)
library("readxl")
library(corrplot)
#install.packages("correlation_matrix")
library(psych)
library(tidyr)
#install.packages("broom")  # For tidying the model summary
#library(broom)
#install.packages("officer")  # For exporting to Word
library(officer)
#install.packages("mice")  # Install the mice package
library(mice)    
#install.packages("naniar")
library(naniar)



```


#load the dataset
```{r}

#full datapath: FINAL_Crypto_DATASET_EEG_CANTAB <- read_excel("Dropbox (BCH)/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/FINAL_Crypto_DATASET_EEG_CANTAB.xlsx")

dat <- FINAL_Crypto_DATASET_EEG_CANTAB <- read_excel("FINAL_Crypto_DATASET_EEG_CANTAB.xlsx")


#make categorical variables factors
dat$cohort <- as.factor(dat$cohort)

dat$sex <- as.factor(dat$sex)
summary(dat$sex)

dat$child_use_screens <- as.factor(dat$child_use_screens)

#make ages numeric
dat$age_mo_EEG_2yr <- as.numeric(dat$age_mo_EEG_2yr)

```

#add new variables needed for revision
#load quality metric assessment 
```{r}
quality_6mo <- HAPPE_quality_metrics_6mo_crypto <- read_csv("HAPPE_quality_metrics_6mo_crypto.csv")

quality_2yr <- HAPPE_quality_metrics_2yr_crypto <- read_csv("HAPPE_quality_metrics_2yr_crypto.csv")

#restrict to just # of segments post-artifact detection
num_segs_6mo <- quality_6mo %>% select(sid, Number_Segs_Post_Seg_Rej_6mo)
num_segs_2yr <- quality_2yr %>% select(sid, Number_Segs_Post_Seg_Rej_2yr)

#merge 6mo and 2yr num seg data
num_segs <- num_segs_6mo %>%
  full_join(num_segs_2yr, by = "sid")

#merge num segs data with final dataset
dat <- merged_data <- dat %>%
  left_join(num_segs, by = "sid")

```

#merge exact age of EEG at 2yrs
```{r}
dat_2yrage <- Age_assessment_EEG_2yr_crypto <- read_excel("Age_assessment_EEG_2yr_crypto.xlsx")
dat_2yrage <- dat_2yrage %>% select(sid, age_mo_EEG_2yr)

#remove old age variable that was rounded to nearest month
dat <- dat %>% select(-age_mo_EEG_2yr)

#merge exact age at 2-yr EEG
dat <- merged_data <- dat %>%
  left_join(dat_2yrage, by = "sid")


```


#log transform EEG variables
```{r}

dat$Log_central_z_T_6mo <- log10(dat$central_z_T_6mo)
dat$Log_frontal_l_T_6mo <- log10(dat$frontal_l_T_6mo)
dat$Log_frontal_r_T_6mo <- log10(dat$frontal_r_T_6mo)

dat$Log_central_z_B_6mo <- log10(dat$central_z_B_6mo)
dat$Log_frontal_l_B_6mo <- log10(dat$frontal_l_B_6mo)
dat$Log_frontal_r_B_6mo <- log10(dat$frontal_r_B_6mo)

dat$Log_central_z_T_2yr <- log10(dat$central_z_T_2yr)
dat$Log_frontal_l_T_2yr <- log10(dat$frontal_l_T_2yr)
dat$Log_frontal_r_T_2yr <- log10(dat$frontal_r_T_2yr)

dat$Log_central_z_B_2yr <- log10(dat$central_z_B_2yr)
dat$Log_frontal_l_B_2yr <- log10(dat$frontal_l_B_2yr)
dat$Log_frontal_r_B_2yr <- log10(dat$frontal_r_B_2yr)

#LOG descriptives
summary(dat$Log_central_z_T_6mo)
#below 3SD would be 0.53597

summary(dat$Log_frontal_l_T_6mo)
#below 3 SD would 1.1976 - 0.52498 = 0.6726



summary(dat$Log_frontal_r_T_6mo)
#below 3SD would be 0.687


summary(dat$Log_central_z_T_2yr)
#below 3SD would be 0.67737

summary(dat$Log_frontal_l_T_2yr)
#below 3SD would be 0.8533239

summary(dat$Log_frontal_r_T_2yr)
#below 3SD would be 0.8019

#remove observations if below 3SD from mean 
below_6mo_central_z_rows <- dat[dat$Log_central_z_T_6mo < 0.53597, ]
print(below_6mo_central_z_rows)

below_6mo_frontal_l_rows <- dat[dat$Log_frontal_l_T_6mo < 0.6726, ]
print(below_6mo_frontal_l_rows)

below_6mo_frontal_r_rows <- dat[dat$Log_frontal_l_T_6mo < 0.687, ]
print(below_6mo_frontal_r_rows)


#2yo - no one to be removed
below_2yr_central_z_rows <- dat[dat$Log_central_z_T_2yr < 0.67737, ]
print(below_2yr_central_z_rows)

below_2yr_frontal_l_rows <- dat[dat$Log_frontal_l_T_2yr < 0.8533239, ]
print(below_2yr_frontal_l_rows)

below_2yr_frontal_r_rows <- dat[dat$Log_frontal_l_T_2yr < 0.8019, ]
print(below_2yr_frontal_r_rows)


#no outliers for 2yr
write_xlsx(dat, "EEG_theta_data_wide.xlsx")





```

#calculate number of families living in poverty
```{r}
dat$income_to_needs_per_day <- dat$income_to_needs / 30.437
summary(dat$income_to_needs_per_day)

dat$extreme_poverty <- cut(dat$income_to_needs_per_day,
                       breaks=c(-Inf, 167.42, Inf),
                       labels=c('poverty', 'not_poverty'))

summary(dat$extreme_poverty)
108/(108+8)


#calculate income per day 
dat$income_per_day <- dat$income / 30.437
summary(dat$income_per_day)


#impute HAZ
dat$HAZ_6mo_impute <- impute(dat$HAZ_6mo, method = "mean")
summary(dat$HAZ_6mo_impute)

#impute income
dat$income_per_day_impute <- impute(dat$income_per_day, method = "mean")




```

#descriptive statistics environmental variables
```{r}

summary(dat$sex)

summary(dat$age_mo_EEG_6mo)
sd(dat$age_mo_EEG_6mo, na.rm=TRUE)
ggplot(dat, aes(x=age_mo_EEG_6mo)) + 
  geom_histogram() 

summary(dat$age_mo_EEG_2yr)
sd(dat$age_mo_EEG_2yr, na.rm=TRUE)
ggplot(dat, aes(x=age_mo_EEG_2yr)) + 
  geom_histogram() 

summary(dat$age_mo_5yr_cog_ass)
sd(dat$age_mo_5yr_cog_ass, na.rm=TRUE)
ggplot(dat, aes(x=age_mo_5yr_cog_ass)) + 
  geom_histogram() 

summary(dat$FCI_6mo)
sd(dat$FCI_6mo, na.rm=TRUE)
ggplot(dat, aes(x=FCI_6mo)) + 
  geom_histogram() 

summary(dat$paternal_yrsofed)
sd(dat$paternal_yrsofed, na.rm=TRUE)
ggplot(dat, aes(x=paternal_yrsofed)) + 
  geom_histogram() 

summary(dat$maternal_yrsofed)
sd(dat$maternal_yrsofed, na.rm=TRUE)
ggplot(dat, aes(x=maternal_yrsofed)) + 
  geom_histogram() 

summary(dat$income)
sd(dat$income, na.rm=TRUE)
ggplot(dat, aes(x=income)) + 
  geom_histogram() 


summary(dat$income_to_needs)
sd(dat$income_to_needs, na.rm=TRUE)
ggplot(dat, aes(x=income_to_needs)) + 
  geom_histogram() 

summary(dat$income_to_needs_per_day)
sd(dat$income_to_needs_per_day, na.rm=TRUE)
ggplot(dat, aes(x=income_to_needs_per_day)) + 
  geom_histogram() 


summary(dat$income_per_day)
sd(dat$income_per_day)

summary(dat$HAZ_6mo)
sd(dat$HAZ_6mo, na.rm=TRUE)
ggplot(dat, aes(x=HAZ_6mo)) + 
  geom_histogram() 

summary(dat$social_support_6mo)
sd(dat$social_support_6mo, na.rm=TRUE)
ggplot(dat, aes(x=social_support_6mo)) + 
  geom_histogram() 

summary(dat$PSS_sum_score)
sd(dat$PSS_sum_score, na.rm=TRUE)
ggplot(dat, aes(x=PSS_sum_score)) + 
  geom_histogram() 

```


#calculate number of stunted children 
```{r}

dat$stunted <- cut(dat$HAZ_6mo,
                       breaks=c(-Inf, -2, Inf),
                       labels=c('stunted', 'not_stunted'))

summary(dat$stunted)
18/(93+18)

```

#Calculate # of mothers with low, medium, high stress 
```{r}
#0-13 = low stress

#14-26 = moderate stress

#27-40 = moderate stress 

dat$PSS_categorites <- cut(dat$PSS_sum_score,
                       breaks=c(0, 13, 26, 40),
                       labels=c('Low', "Moderate", 'High'))

summary(dat$PSS_categorites)
22/115
78/115
15/115

```


#Theta EEG descriptives
```{r}
summary(dat$central_z_T_6mo)
sd(dat$central_z_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=central_z_T_6mo)) + 
  geom_histogram() 


summary(dat$frontal_l_T_6mo)
sd(dat$frontal_l_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=frontal_l_T_6mo)) + 
  geom_histogram() 

summary(dat$frontal_r_T_6mo)
sd(dat$frontal_r_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=frontal_r_T_6mo)) + 
  geom_histogram() 


summary(dat$central_z_T_2yr)
sd(dat$central_z_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=central_z_T_2yr)) + 
  geom_histogram() 

summary(dat$frontal_l_T_2yr)
sd(dat$frontal_l_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=frontal_l_T_2yr)) + 
  geom_histogram() 

summary(dat$frontal_r_T_2yr)
sd(dat$frontal_r_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=frontal_r_T_2yr)) + 
  geom_histogram() 


#LOG descriptives
summary(dat$Log_central_z_T_6mo)
sd(dat$Log_central_z_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=Log_central_z_T_6mo)) + 
  geom_histogram() 
#below 3SD would be 0.53597

summary(dat$Log_frontal_l_T_6mo)
sd = sd(dat$Log_frontal_l_T_6mo, na.rm=TRUE)
3*sd
ggplot(dat, aes(x=Log_frontal_l_T_6mo)) + 
  geom_histogram() 
#below 3 SD would 1.1976 - 0.52498 = 0.6726



summary(dat$Log_frontal_r_T_6mo)
sd(dat$Log_frontal_r_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=Log_frontal_r_T_6mo)) + 
  geom_histogram() 
#below 3SD would be 0.687


summary(dat$Log_central_z_T_2yr)
sd(dat$Log_central_z_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=Log_central_z_T_2yr)) + 
  geom_histogram() 
#below 3SD would be 0.67737

summary(dat$Log_frontal_l_T_2yr)
sd(dat$Log_frontal_l_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=Log_frontal_l_T_2yr)) + 
  geom_histogram() 
#below 3SD would be 0.8533239

summary(dat$Log_frontal_r_T_2yr)
sd(dat$Log_frontal_r_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=Log_frontal_r_T_2yr)) + 
  geom_histogram() 
#below 3SD would be 0.8019



```



#Screen use descriptives
```{r}

summary(dat$child_use_screens)

summary(dat$minutes_on_screens)
sd(dat$minutes_on_screens, na.rm=TRUE)
ggplot(dat, aes(x=minutes_on_screens)) + 
  geom_histogram() 


```

#CANTAB descriptive statistics
```{r}

summary(dat$SWMTE)
ggplot(dat, aes(x=SWMTE)) + 
  geom_histogram() 

summary(dat$PALTEA)
sd(dat$PALTEA, na.rm=TRUE)
ggplot(dat, aes(x=PALTEA)) + 
  geom_histogram() 

summary(dat$SSPFTUE)
ggplot(dat, aes(x=SSPFTUE)) + 
  geom_histogram() 

summary(dat$SOCPSMMT)
ggplot(dat, aes(x=SOCPSMMT)) + 
  geom_histogram() 

```
#GNG descriptive statistics
```{r}
summary(dat$GNG_RT_avg)

ggplot(dat, aes(x=GNG_RT_avg)) + 
  geom_histogram() 

summary(dat$GNG_accuracy_all)
sd(dat$GNG_accuracy_all, na.rm=TRUE)

ggplot(dat, aes(x=GNG_accuracy_all)) + 
  geom_histogram() 


```


#WPPSI descriptive statistics
```{r}
summary(dat$viq_5yr)
ggplot(dat, aes(x=viq_5yr)) + 
  geom_histogram() 

summary(dat$piq_5yr)
ggplot(dat, aes(x=piq_5yr)) + 
  geom_histogram() 

summary(dat$fsiq_5yr)
sd(dat$fsiq_5yr, na.rm=TRUE)
ggplot(dat, aes(x=fsiq_5yr)) + 
  geom_histogram() 

cor.test(dat$fsiq_5yr, dat$piq_5yr) #0.89 correlation


```

#Correlation between screen use and cantab?
```{r}
cor.test(dat$minutes_on_screens, dat$PALTEA)
cor.test(dat$minutes_on_screens, dat$SWMTE)
cor.test(dat$minutes_on_screens, dat$SSPFTUE)
cor.test(dat$minutes_on_screens, dat$SOCPSMMT)

cor.test(dat$minutes_on_screens, dat$GNG_accuracy_go)


##no correlation between minutes on screen and any CANTAB outcome

```

#correlations between maternal stress, depression, and social support
```{r}
cor.test(dat$social_support_6mo, dat$EPDS_6mo)
cor.test(dat$PSS_sum_score, dat$EPDS_6mo)
cor.test(dat$PSS_sum_score, dat$social_support_6mo)


```




#Correlation matrices of non-EEG measures 
```{r}
#correlation matrix with everything possible 
corr_matrix_adv_beh <- select (dat, c(age_mo_EEG_6mo, age_mo_EEG_2yr, age_mo_5yr_cog_ass, maternal_yrsofed, paternal_yrsofed, income_to_needs_per_day, HAZ_6mo, FCI_6mo, PSS_sum_score, social_support_6mo, EPDS_6mo, minutes_on_screens, SWMTE, PALTEA, SSPFTUE, SOCPSMMT, viq_5yr, piq_5yr, fsiq_5yr))

#colnames(corr_matrix_NIRS_2y) <- c("RSTG High TP-Novel", "LSTG High TP-Novel", "Child Adversity", "Maternal Adversity", "Environmental Adversity", "Income-to-Needs", "Maternal Education", "HAZ", "Language (Mullen)")


rcorr(as.matrix(corr_matrix_adv_beh), type = c("pearson"))
pval <- corr.test(corr_matrix_adv_beh, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_adv_beh <- cor(corr_matrix_adv_beh, use="pairwise.complete.obs")

corrplot(corr_adv_beh, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")


#correlation matrix with just the selected variables for final models 

#covariates selected - age, sex, income to needs, maternal education
#adversity measures selected - HAZ, FCI
#CANTAB measure selected - PALTEA
#WPPSI measure selected - FSIQ

#correlation matrix with selected variables
corr_matrix_select_adv_beh <- select (dat, c(maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, GNG_accuracy_all, PALTEA,fsiq_5yr))

rcorr(as.matrix(corr_matrix_select_adv_beh), type = c("pearson"))
pval <- corr.test(corr_matrix_select_adv_beh, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_select_adv_beh <- cor(corr_matrix_select_adv_beh, use="pairwise.complete.obs")

corrplot(corr_select_adv_beh, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")



```




#check whether line noise correlated with power
```{r}

#correlation matrix with selected variables
corr_matrix_theta_linenoise <- select (dat, c(central_z_T_6mo, frontal_l_T_6mo, frontal_r_T_6mo, central_z_T_2yr, frontal_l_T_2yr, frontal_r_T_2yr, central_z_B_6mo, frontal_l_B_6mo, frontal_r_B_6mo, central_z_B_2yr, frontal_l_B_2yr, frontal_r_B_2yr, linenoise_reduction_50hz_6mo, linenoise_reduction_50hz_2yr))

corr_matrix_log_theta_linenoise <- select (dat, c(Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, linenoise_reduction_50hz_6mo, linenoise_reduction_50hz_2yr))
corr_matrix_log_theta_linenoise

rcorr(as.matrix(corr_matrix_log_theta_linenoise), type = c("pearson"))
pval <- corr.test(corr_matrix_log_theta_linenoise, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_theta_linenoise <- cor(corr_matrix_log_theta_linenoise, use="pairwise.complete.obs")

corrplot(corr_theta_linenoise, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")

```


#corr matrix with all included variables (log EEG)
```{r}
#correlation matrix with selected variables
corr_matrix_EEG_behavior <- select (dat, c(Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, PALTEA, GNG_accuracy_all, fsiq_5yr, Number_Segs_Post_Seg_Rej_6mo, Number_Segs_Post_Seg_Rej_2yr))

new_column_names <- c("6mo Frontocentral Theta", "6mo Frontal Left Theta", "6mo Frontal Right Theta", "2yr Frontocentral Theta", "2yr Frontal Left Theta", "2yr Frontal Right Theta", "Maternal Education (yrs)", "Adjusted Family Income", "HAZ", "Maternal Stress", "Working Memory", "Inhibition", "Full-Scale IQ", "6mo # Included Segments", "2yr # Included Segments")

colnames(corr_matrix_EEG_behavior) <- new_column_names

rcorr(as.matrix(corr_matrix_EEG_behavior), type = c("pearson"))
pval <- corr.test(corr_matrix_EEG_behavior, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_EEG_behavior <- cor(corr_matrix_EEG_behavior, use="pairwise.complete.obs")

corrplot(corr_EEG_behavior, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")

#try to make it so stars and numbers appear at the same time
#trace(corrplot, edit=TRUE)



corrplot(corr_EEG_behavior, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', type="lower", diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.8, number.cex=0.4, addCoef.col ='black',
         insig = 'label_sig', pch.col = 'black', order = 'original', tl.col="black")






```

#corr matrix with change in theta power 
```{r}
#change in theta power variables
dat$Log_central_z_T_change <- dat$Log_central_z_T_2yr - dat$Log_central_z_T_6mo

dat$Log_frontal_l_T_change <- dat$Log_frontal_l_T_2yr - dat$Log_frontal_l_T_6mo

dat$Log_frontal_r_T_change <- dat$Log_frontal_r_T_2yr - dat$Log_frontal_r_T_6mo



corr_matrix_log_theta_change <- select (dat, c(Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, Log_central_z_T_change, Log_frontal_l_T_change, Log_frontal_r_T_change, maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, PALTEA, GNG_accuracy_all, fsiq_5yr))

rcorr(as.matrix(corr_matrix_log_theta_change), type = c("pearson"))
pval <- corr.test(corr_matrix_log_theta_change, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_EEG_behavior <- cor(corr_matrix_log_theta_change, use="pairwise.complete.obs")

corrplot(corr_matrix_log_theta_change, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")




```

#Aim 1 - LMMs for just theta power 

#create long dataset (convert from wide)
```{r}

#create newdataset with only theta power and rename columns 

dat_EEG_sub <- select (dat, c(sid, Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, sex, maternal_yrsofed, income_per_day_impute, HAZ_6mo_impute, PSS_sum_score, age_mo_EEG_6mo, age_mo_EEG_2yr, Number_Segs_Post_Seg_Rej_6mo, Number_Segs_Post_Seg_Rej_2yr))

#rename columns 
colnames(dat_EEG_sub) <- c("sid", "6mo_CentralZ", "6mo_FrontalL", "6mo_FrontalR", "2yr_CentralZ", "2yr_FrontalL", "2yr_FrontalR", "sex", "maternal_yrsofed", "income_per_day_impute", "HAZ_6mo_impute", "PSS_sum_score", "age_mo_EEG_6mo", "age_mo_EEG_2yr", "num_segs_6mo", "num_segs_2yr")

library(tidyr)

dat_EEG_long <- pivot_longer(dat_EEG_sub, 
                          cols = -c(sid, sex, maternal_yrsofed, income_per_day_impute, HAZ_6mo_impute, PSS_sum_score, age_mo_EEG_6mo, age_mo_EEG_2yr, num_segs_6mo, num_segs_2yr), 
                          names_to = c("timepoint", "ROI"), 
                          names_sep = "_", 
                          values_to = "theta_power"
) %>%
  # Create age_month_EEG based on the timepoint
mutate(
    age_mo_EEG = case_when(
      timepoint == "6mo" ~ age_mo_EEG_6mo,
      timepoint == "2yr" ~ age_mo_EEG_2yr
    ),
    # Create num_segs based on the timepoint
    num_segs_EEG = case_when(
      timepoint == "6mo" ~ num_segs_6mo,
      timepoint == "2yr" ~ num_segs_2yr
    )
  )


#make 6mo the reference category
dat_EEG_long$timepoint <- factor(dat_EEG_long$timepoint)
summary(dat_EEG_long$timepoint)
dat_EEG_long$timepoint <- relevel(dat_EEG_long$timepoint, ref = "6mo")


#now need to create three separate datasets for each ROI separately 
dat_EEG_long_central_z <- filter(dat_EEG_long, ROI == "CentralZ")
dat_EEG_long_frontal_l <- filter(dat_EEG_long, ROI == "FrontalL")
dat_EEG_long_frontal_r <- filter(dat_EEG_long, ROI == "FrontalR")

#center age_mo_EEG variable so centered at 6mo

```


#Model 1: one big model with just EEG data
```{r}
#random intercept for participant - can have different baseline levels of theta
#random slope for participant - can have different slopes between 6mo and 2yr
lmm_rand_time <- lmer(theta_power ~ ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long)

summary(lmm_rand_time)
confint(lmm_rand_time)
AIC(lmm_rand_time)
BIC(lmm_rand_time)
logLik(lmm_rand_time)


#variance explained by individual variation: 
0.024093 / (0.024093 + 0.037238 + 0.004808)
#variance explained by timepoint:
0.037238 / (0.024093 + 0.037238 + 0.004808)
#variance explained by residual:
0.004808 / (0.024093 + 0.037238 + 0.004808)

#model 1 with number of segements added
lmm_rand_time <- lmer(theta_power ~ ROI * timepoint + num_segs_EEG + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time)



#do this model with age instead of timepoint
lmm_rand_time <- lmer(theta_power ~ ROI * age_mo_EEG + (1 + age_mo_EEG| sid), data = dat_EEG_long)

summary(lmm_rand_time)
confint(lmm_rand_time)
AIC(lmm_rand_time)
BIC(lmm_rand_time)
logLik(lmm_rand_time)

#variance explained by individual variation: 
0.0364739 / (0.0364739 + 0.0000821 + 0.0047808)
#variance explained by age:
0.0000821 / (0.0364739 + 0.0000821 + 0.0047808)
#variance explained by residual:
0.0047808 / (0.0364739 + 0.0000821 + 0.0047808)




```


#Aim 2: add adversity measures to one big model
```{r}

#impute HAZ and income (above)

#no interaction terms
lmm_rand_time_adv <- lmer(theta_power ~ sex + maternal_yrsofed + income_per_day_impute + HAZ_6mo_impute + PSS_sum_score + ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time_adv)
confint(lmm_rand_time_adv)
AIC(lmm_rand_time_adv)
BIC(lmm_rand_time_adv)
logLik(lmm_rand_time_adv)

summary(dat$HAZ_6mo_impute)

#no interaction terms with num segments EEG included as fixed effect
lmm_rand_time_adv <- lmer(theta_power ~ sex + maternal_yrsofed + income_per_day_impute + HAZ_6mo_impute + PSS_sum_score  + num_segs_EEG + ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time_adv)
confint(lmm_rand_time_adv)
AIC(lmm_rand_time_adv)
BIC(lmm_rand_time_adv)
logLik(lmm_rand_time_adv)

summary(dat$HAZ_6mo_impute)

#interaction terms
lmm_rand_time_adv_inter <- lmer(theta_power ~ sex*timepoint + maternal_yrsofed*timepoint + income_per_day_impute*timepoint + HAZ_6mo_impute*timepoint + PSS_sum_score*timepoint + ROI*timepoint + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time_adv_inter)
confint(lmm_rand_time_adv_inter)
nobs(lmm_rand_time_adv_inter)

AIC(lmm_rand_time_adv_inter)
BIC(lmm_rand_time_adv_inter)
logLik(lmm_rand_time_adv_inter)

#interaction terms with number EEG segs as fixed effect
lmm_rand_time_adv_inter <- lmer(theta_power ~ num_segs_EEG + sex*timepoint + maternal_yrsofed*timepoint + income_per_day_impute*timepoint + HAZ_6mo_impute*timepoint + PSS_sum_score*timepoint + ROI*timepoint + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time_adv_inter)
confint(lmm_rand_time_adv_inter)
nobs(lmm_rand_time_adv_inter)

AIC(lmm_rand_time_adv_inter)
BIC(lmm_rand_time_adv_inter)
logLik(lmm_rand_time_adv_inter)


#worse AIC and BIC when include adversity measures - does not improve the model



```


#Likelihood ratio tests 
```{r}

anova(lmm_rand_time, lmm_rand_time_adv)

anova(lmm_rand_time_adv, lmm_rand_time_adv_inter)

anova(lmm_rand_time, lmm_rand_time_adv_inter)



```


#t-tests for ROI differences within age (probably don't need this)
```{r}

#CREATE AVERAGE OF THETA POWER 
dat$Log_T_6mo_avg <- (dat$Log_central_z_T_6mo + dat$Log_frontal_l_T_6mo + dat$Log_frontal_r_T_6mo)/3
dat$Log_T_2yr_avg <- (dat$Log_central_z_T_2yr + dat$Log_frontal_l_T_2yr + dat$Log_frontal_r_T_2yr)/3

t.test(dat$Log_T_6mo_avg, dat$Log_T_2yr_avg)



#do follow-up t-tests to see whether differences between ROIs within each timepoint

t.test(dat$Log_central_z_T_6mo, dat$Log_frontal_l_T_6mo)
t.test(dat$Log_central_z_T_6mo, dat$Log_frontal_r_T_6mo)
t.test(dat$Log_frontal_l_T_6mo, dat$Log_frontal_r_T_6mo)
#difference between centralZ and two frontal ones (higher in frontal), but no difference between frontal l and r... more lateralized (but not to one side or the other)

t.test(dat$Log_central_z_T_2yr, dat$Log_frontal_l_T_2yr)
t.test(dat$Log_central_z_T_2yr, dat$Log_frontal_r_T_2yr)
t.test(dat$Log_frontal_l_T_2yr, dat$Log_frontal_r_T_2yr)
#difference between frontal l and central z (higher in frontal l) at least before FDR. no difference between central z and frontal r or frontal l or r

#conclusion - maybe more evenly distributed at 2yrs 


#need to correct for multiple comparisons if go this route (correct for 3 or six tests?)

#correcting just within 6mo
p_values_6mo <- c(0.01218509, 0.0141507, 0.910078)
fdr_corrected_p_values_6mo <- p.adjust(p_values_6mo, method = "fdr")
fdr_corrected_p_values_6mo


#correcting just within 2yr
p_values_2yr <- c(0.03264265, 0.189457, 0.3962806)
fdr_corrected_p_values_2yr <- p.adjust(p_values_2yr, method = "fdr")
fdr_corrected_p_values_2yr


#correcting within everything
p_values_all <- c(0.01218509, 0.0141507, 0.910078, 0.03264265, 0.189457, 0.3962806)
fdr_corrected_p_values_all <- p.adjust(p_values_all, method = "fdr")
fdr_corrected_p_values_all



```

#create a bar plot to display these results
```{r}
#relevel ROI
dat_EEG_long$ROI_relev <- factor(dat_EEG_long$ROI, levels = c("FrontalL", "CentralZ", "FrontalR"))

#calculate standard error of each group
summary_data <- dat_EEG_long %>%
  group_by(timepoint, ROI_relev) %>%
  summarise(mean_theta_power = mean(theta_power, na.rm=TRUE),
            se_theta_power = sd(theta_power, na.rm=TRUE) / sqrt(n()))

ggplot(summary_data, aes(x = timepoint, y = mean_theta_power, fill = ROI_relev)) +
  geom_bar(stat = "identity", position = "dodge") + 
geom_errorbar(aes(ymin = mean_theta_power - se_theta_power, ymax = mean_theta_power + se_theta_power),
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(x = "Timepoint", y = "Log Absolute Theta Power") +
  scale_fill_manual(values = c("FrontalL" = "orange", "CentralZ" = "purple", "FrontalR" = "green3"), 
                    labels = c("Frontal Left", "Frontocentral", "Frontal Right")) +
  coord_cartesian(ylim = c(1, 1.4)) +
  theme_minimal() + 
  theme(panel.grid = element_blank()) + 
  labs(fill = "ROI")



ggplot(dat_EEG_long, aes(x = timepoint, y = theta_power, fill = ROI_relev)) +
  geom_bar(stat = "summary", position = "dodge") +
  labs(x = "Timepoint", y = "Log Absolute Theta Power") +
  scale_fill_manual(values = c("FrontalL" = "orange", "CentralZ" = "purple", "FrontalR" = "green3"), 
                    labels = c("Frontal Left", "Frontocentral", "Frontal Right")) +
  theme_minimal() + 
  theme(panel.grid = element_blank()) + 
  labs(fill = "ROI") + 
      coord_cartesian(ylim = c(1, 1.75)) # Adjust y-axis limits without affecting the bars 


  


```



#Aim 3 - regressions predicting cognition

#Regressions: adversity => behavior
```{r}

adv_WPPSI <- lm (fsiq_5yr ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score, data = dat)
summary(adv_WPPSI)
 nobs(adv_WPPSI)
 
 adv_PALTEA <- lm (PALTEA ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score, data = dat)
summary(adv_PALTEA)
 nobs(adv_PALTEA)


adv_GNG <- lm (GNG_accuracy_all ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score + fsiq_5yr, data = dat)
summary(adv_GNG)
nobs(adv_GNG)


adv_PALTEA_control_IQ <- lm (PALTEA ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score + fsiq_5yr, data = dat)
summary(adv_PALTEA_control_IQ)
 nobs(adv_PALTEA)


adv_GNG_control_IQ <- lm (GNG_accuracy_all ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score + fsiq_5yr, data = dat)
summary(adv_GNG_control_IQ)
nobs(adv_GNG)


 #correct for multiple comparisons (FDR)
 p_values_adv_WPPSI <- summary(adv_WPPSI)$coef[, "Pr(>|t|)"]
 p_values_adv_WPPSI
p_values_adv_PALTEA <- summary(adv_PALTEA)$coef[, "Pr(>|t|)"]
p_values_adv_PALTEA
p_values_adv_GNG <- summary(adv_GNG)$coef[, "Pr(>|t|)"]

all_p_values_adv <- c(p_values_adv_WPPSI, p_values_adv_PALTEA, p_values_adv_GNG)
all_p_values_adv
corrected_p_values_adv <- p.adjust(all_p_values_adv, method = "fdr")
corrected_p_values_adv


```

#Regressions: 2yr EEG => Cognition

```{r}
options(digits=10)
theta_WPPSI <- lm(fsiq_5yr ~ Log_central_z_T_2yr + Log_frontal_l_T_2yr + Log_frontal_r_T_2yr, data = dat)
summary(theta_WPPSI)
nobs(theta_WPPSI)

theta_GNG <- lm(GNG_accuracy_all ~ Log_central_z_T_2yr + Log_frontal_l_T_2yr + Log_frontal_r_T_2yr, data = dat)
summary(theta_GNG)
nobs(theta_GNG)


theta_PALTEA <- lm(PALTEA ~ Log_central_z_T_2yr + Log_frontal_l_T_2yr + Log_frontal_r_T_2yr, data = dat)
summary(theta_PALTEA)
nobs(theta_PALTEA)


 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI <- summary(theta_WPPSI)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI
p_values_theta_PALTEA <- summary(theta_PALTEA)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA
p_values_theta_GNG <- summary(theta_GNG)$coef[, "Pr(>|t|)"]
p_values_theta_GNG

all_p_values_theta <- c(p_values_theta_WPPSI, p_values_theta_PALTEA, p_values_theta_GNG)
all_p_values_theta
corrected_p_values_theta <- p.adjust(all_p_values_theta, method = "fdr")
corrected_p_values_theta

```


#Regressions: 6mo EEG => Cognition

```{r}
options(digits=10)
theta_WPPSI <- lm(fsiq_5yr ~ Log_central_z_T_6mo + Log_frontal_l_T_6mo + Log_frontal_r_T_6mo, data = dat)
summary(theta_WPPSI)
nobs(theta_WPPSI)

theta_GNG <- lm(GNG_accuracy_all ~ Log_central_z_T_6mo + Log_frontal_l_T_6mo + Log_frontal_r_T_6mo, data = dat)
summary(theta_GNG)
nobs(theta_GNG)


theta_PALTEA <- lm(PALTEA ~ Log_central_z_T_6mo + Log_frontal_l_T_6mo + Log_frontal_r_T_6mo, data = dat)
summary(theta_PALTEA)
nobs(theta_PALTEA)


 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI <- summary(theta_WPPSI)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI
p_values_theta_PALTEA <- summary(theta_PALTEA)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA
p_values_theta_GNG <- summary(theta_GNG)$coef[, "Pr(>|t|)"]
p_values_theta_GNG

all_p_values_theta <- c(p_values_theta_WPPSI, p_values_theta_PALTEA, p_values_theta_GNG)
all_p_values_theta
corrected_p_values_theta <- p.adjust(all_p_values_theta, method = "fdr")
corrected_p_values_theta

```


#mediation analyses
```{r}
#HAZ => Central_Z theta => FSIQ

#open and run process.R file
#IV to mediator = a path
#mediator to DV = b path (Central Z to IQ)
#direct path IV to DV = c' path (HAZ => IQ)
#total effect = c path


#with just central_z
process (data = dat, y = "fsiq_5yr", x = "HAZ_6mo", m ="central_z_T_2yr", model = 4)
process (data = dat, y = "fsiq_5yr", x = "HAZ_6mo", m ="central_z_T_2yr", model = 4, effsize =1, total =1, stand =1, contrast =1, boot = 10000 , modelbt = 1, seed = 654321)


#with all three ROIs
process(data = dat, y = "fsiq_5yr", x = "HAZ_6mo", m =c("central_z_T_2yr", "frontal_l_T_2yr", "frontal_r_T_2yr"), model = 4, effsize =1, total =1, stand =1, contrast =1, boot = 10000 , modelbt = 1, seed = 654321)


```


#REVISION 1 Child Development

#load dataset with all 130 SIDs for sensitivity analysis
```{r}
dat_full <- Full_Sample_Demographics_For_Sensitivity_Analysis <- read_excel("Full_Sample_Demographics_For_Sensitivity_Analysis.xlsx")

# Create the boxplot for paternal ed
ggplot(dat_full, aes(x = Inclusion, y = paternal_ed_yrs)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Paternal Education Years by Inclusion Status",
    x = "Inclusion Status",
    y = "Paternal Education (Years)"
  ) +
  theme_minimal()

t.test(dat_full$paternal_ed_yrs ~ dat_full$Inclusion)

# Create the boxplot for maternal ed
ggplot(dat_full, aes(x = Inclusion, y = maternal_ed_yrs)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Maternal Education Years by Inclusion Status",
    x = "Inclusion Status",
    y = "Maternal Education (Years)"
  ) +
  theme_minimal()

t.test(dat_full$maternal_ed_yrs ~ dat_full$Inclusion)

# Create the boxplot for income adjusted
ggplot(dat_full, aes(x = Inclusion, y = income_to_needs)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Adjusted Income by Inclusion Status",
    x = "Inclusion Status",
    y = "Adjusted Income"
  ) +
  theme_minimal()

t.test(dat_full$income_to_needs ~ dat_full$Inclusion)


# Create the boxplot for PSS 
ggplot(dat_full, aes(x = Inclusion, y = PSS_sum_score)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Perceived Stress Scale",
    x = "Inclusion Status",
    y = "Perceived Stress Scale Score by Inclusion Status"
  ) +
  theme_minimal()

t.test(dat_full$PSS_sum_score ~ dat_full$Inclusion)


# Create the boxplot for HAZ 
ggplot(dat_full, aes(x = Inclusion, y = HAZ_6mo)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of HAZ by Inclusion Status",
    x = "Inclusion Status",
    y = "HAZ"
  ) +
  theme_minimal()

t.test(dat_full$HAZ_6mo ~ dat_full$Inclusion)



summary_stats <- dat_full %>%
  group_by(Inclusion) %>%
  summarise(
    mean_paternal_ed_yrs = mean(paternal_ed_yrs, na.rm = TRUE),
    sd_paternal_ed_yrs = sd(paternal_ed_yrs, na.rm = TRUE),
    median_paternal_ed_yrs = median(paternal_ed_yrs, na.rm = TRUE),
    min_paternal_ed_yrs = min(paternal_ed_yrs, na.rm = TRUE),
    max_paternal_ed_yrs = max(paternal_ed_yrs, na.rm = TRUE),
    n = n()
  )
summary_stats



```


#Test for MCAR
```{r}
dat_mcar_test <- dat %>% select(Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, PALTEA, GNG_accuracy_all, fsiq_5yr)
str(dat_mcar_test)

gg_miss_var(dat_mcar_test)1

mcar_test_result <- naniar::mcar_test(dat_mcar_test)
print(mcar_test_result)


```
#Relative power calculations

#calculation and merging with long dataset
```{r}
#load datasheet with all bands power 
dat_all_bands <- All_bands_power_6mo_2yr <- read_excel("All_bands_power_6mo_2yr.xlsx")

#calculate relative power 
dat_all_bands$central_z_relativeT <- dat_all_bands$Theta_CentralZ / (dat_all_bands$Delta_CentralZ + dat_all_bands$Theta_CentralZ + dat_all_bands$LowAlpha_CentralZ + dat_all_bands$HighAlpha_CentralZ + dat_all_bands$Beta_CentralZ + dat_all_bands$Gamma_CentralZ)
summary(dat_all_bands$central_z_relativeT)

dat_all_bands$frontal_l_relativeT <- dat_all_bands$Theta_FrontalL / (dat_all_bands$Delta_FrontalL + dat_all_bands$Theta_FrontalL + dat_all_bands$LowAlpha_FrontalL + dat_all_bands$HighAlpha_FrontalL + dat_all_bands$Beta_FrontalL + dat_all_bands$Gamma_FrontalL)


dat_all_bands$frontal_r_relativeT <- dat_all_bands$Theta_FrontalR / (dat_all_bands$Delta_FrontalR + dat_all_bands$Theta_FrontalR + dat_all_bands$LowAlpha_FrontalR + dat_all_bands$HighAlpha_FrontalR + dat_all_bands$Beta_FrontalR + dat_all_bands$Gamma_FrontalR)


#restrict dataset to just relative power 
dat_rel_theta <- subset(dat_all_bands, select = c(sid, timepoint, central_z_relativeT, frontal_l_relativeT, frontal_r_relativeT))

#convert to long format so matches dat_EEG_long
colnames(dat_rel_theta) <- c("sid", "timepoint", "relativeT_CentralZ", "relativeT_FrontalL", "relativeT_FrontalR")


library(tidyr)

dat_rel_theta_long <- pivot_longer(dat_rel_theta, 
                          cols = -c(sid, timepoint), 
                          names_to = c("ROI"), 
                          values_to = "relative_theta_power")

dat_rel_theta_long$ROI <- gsub("^relativeT_", "", dat_rel_theta_long$ROI)

#merge relative theta data with dat_EEG_long
dat_EEG_long_with_rel <- dat_EEG_long %>%
  left_join(dat_rel_theta_long, by = c("sid", "timepoint", "ROI"))


#make 6mo the reference category
dat_EEG_long_with_rel$timepoint <- factor(dat_EEG_long_with_rel$timepoint)
summary(dat_EEG_long_with_rel$timepoint)
dat_EEG_long_with_rel$timepoint <- relevel(dat_EEG_long_with_rel$timepoint, ref = "6mo")
summary(dat_EEG_long_with_rel$timepoint)

  
```

#run models with relative power
```{r}
#Model 1
#random intercept for participant - can have different baseline levels of theta
#random slope for participant - can have different slopes between 6mo and 2yr
#with relative power
lmm_rand_time_rel <- lmer(relative_theta_power ~ ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long_with_rel)

summary(lmm_rand_time_rel)
confint(lmm_rand_time_rel)
AIC(lmm_rand_time_rel)
BIC(lmm_rand_time_rel)
logLik(lmm_rand_time_rel)


#Model 2
#no interaction terms with num segments EEG included as fixed effect - with relative power
options(scipen = 999)
lmm_rand_time_adv_rel <- lmer(relative_theta_power ~ sex + maternal_yrsofed + income_per_day_impute + HAZ_6mo_impute + PSS_sum_score  + num_segs_EEG + ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long_with_rel)
summary(lmm_rand_time_adv_rel)
confint(lmm_rand_time_adv_rel)
AIC(lmm_rand_time_adv_rel)
BIC(lmm_rand_time_adv_rel)
logLik(lmm_rand_time_adv_rel)



#Model 3
#interaction terms with number EEG segs as fixed effect - with relative power
lmm_rand_time_adv_inter_rel <- lmer(relative_theta_power ~ num_segs_EEG + sex*timepoint + maternal_yrsofed*timepoint + income_per_day_impute*timepoint + HAZ_6mo_impute*timepoint + PSS_sum_score*timepoint + ROI*timepoint + (1 + timepoint| sid), data = dat_EEG_long_with_rel)
summary(lmm_rand_time_adv_inter_rel)
confint(lmm_rand_time_adv_inter_rel)
nobs(lmm_rand_time_adv_inter_rel)

AIC(lmm_rand_time_adv_inter_rel)
BIC(lmm_rand_time_adv_inter_rel)
logLik(lmm_rand_time_adv_inter_rel)





```
#Other ROI analyses
```{r}
#calculate relative power
dat_otherROIs <- All_bands_ALL_ROIS_power_6mo_2yr <- read_excel("All_bands_ALL_ROIS_power_6mo_2yr.xlsx")

dat_otherROIs$Central_relativeT <- dat_otherROIs$Theta_Central / (dat_otherROIs$Delta_Central + dat_otherROIs$Theta_Central + dat_otherROIs$LowAlpha_Central + dat_otherROIs$HighAlpha_Central + dat_otherROIs$Beta_Central + dat_otherROIs$Gamma_Central)

dat_otherROIs$Frontal_relativeT <- dat_otherROIs$Theta_Frontal / (dat_otherROIs$Delta_Frontal + dat_otherROIs$Theta_Frontal + dat_otherROIs$LowAlpha_Frontal + dat_otherROIs$HighAlpha_Frontal + dat_otherROIs$Beta_Frontal + dat_otherROIs$Gamma_Frontal)

dat_otherROIs$Occipital_relativeT <- dat_otherROIs$Theta_Occipital / (dat_otherROIs$Delta_Occipital + dat_otherROIs$Theta_Occipital + dat_otherROIs$LowAlpha_Occipital + dat_otherROIs$HighAlpha_Occipital + dat_otherROIs$Beta_Occipital + dat_otherROIs$Gamma_Occipital)

dat_otherROIs$Parietal_relativeT <- dat_otherROIs$Theta_Parietal / (dat_otherROIs$Delta_Parietal + dat_otherROIs$Theta_Parietal + dat_otherROIs$LowAlpha_Parietal + dat_otherROIs$HighAlpha_Parietal + dat_otherROIs$Beta_Parietal + dat_otherROIs$Gamma_Parietal)

dat_otherROIs$Temporal_relativeT <- dat_otherROIs$Theta_Temporal / (dat_otherROIs$Delta_Temporal + dat_otherROIs$Theta_Temporal + dat_otherROIs$LowAlpha_Temporal + dat_otherROIs$HighAlpha_Temporal + dat_otherROIs$Beta_Temporal + dat_otherROIs$Gamma_Temporal)

#rename absolute power variables
dat_otherROIs$Central_absoluteT <- log10(dat_otherROIs$Theta_Central)

dat_otherROIs$Frontal_absoluteT <- log10(dat_otherROIs$Theta_Frontal)

dat_otherROIs$Occipital_absoluteT <- log10(dat_otherROIs$Theta_Occipital)

dat_otherROIs$Parietal_absoluteT <- log10(dat_otherROIs$Theta_Parietal)

dat_otherROIs$Temporal_absoluteT <- log10(dat_otherROIs$Theta_Temporal)




#create subset with just theta power (absolute and relative9)
dat_otherROIs <- subset(dat_otherROIs, select = c(sid, timepoint, Central_absoluteT, Frontal_absoluteT, Occipital_absoluteT, Parietal_absoluteT, Temporal_absoluteT, Central_relativeT, Frontal_relativeT, Occipital_relativeT, Parietal_relativeT, Temporal_relativeT))

#pivot longer
# Pivoting longer
dat_otherROIs_long <- dat_otherROIs %>%
  pivot_longer(
    cols = -c(sid, timepoint), 
    names_to = c("otherROIs", "ThetaType"), 
    names_pattern = "(.*)_(.*)",
    values_to = "ThetaValue"
  ) %>%
  pivot_wider(
    names_from = ThetaType, 
    values_from = ThetaValue,
    names_prefix = "Theta_"
  )

#reorder timepoint so 6mo is baseline
dat_otherROIs_long$timepoint <- factor(dat_otherROIs_long$timepoint)
dat_otherROIs_long$timepoint <- relevel(dat_otherROIs_long$timepoint, ref = "6mo")
summary(dat_EEG_long_with_rel$timepoint)

#subset dat_EEG_long so just controls not theta or ROIs
dat_long_no_EEG <- subset(dat_EEG_long, select = -c(ROI, theta_power, num_segs_EEG, timepoint, age_mo_EEG))

dat_wide_no_EEG <- dat_long_no_EEG %>%
  distinct(sid, .keep_all = TRUE)

dat_other_ROIs_long_adv <- dat_otherROIs_long %>%
  left_join(dat_wide_no_EEG, by = "sid")

# Create a single num_segs column based on timepoint
dat_other_ROIs_long_adv2 <- dat_other_ROIs_long_adv %>%
  mutate(
    num_segs_EEG = case_when(
      timepoint == "6mo" ~ num_segs_6mo,
      timepoint == "2yr" ~ num_segs_2yr
    )
  ) %>%
  # Drop the original num_segs_6mo and num_segs_2yr columns
  select(-num_segs_6mo, -num_segs_2yr)

# Create a single age_EEG column based on timepoint
dat_other_ROIs_long_adv2 <- dat_other_ROIs_long_adv2 %>%
  mutate(
    age_mo_EEG = case_when(
      timepoint == "6mo" ~ age_mo_EEG_6mo,
      timepoint == "2yr" ~ age_mo_EEG_6mo
    )
  ) %>%
  # Drop the original num_segs_6mo and num_segs_2yr columns
  select(-age_mo_EEG_6mo, -age_mo_EEG_2yr)

dat_otherRegions <- dat_other_ROIs_long_adv2

```

#Modeling with other regions: absolute power
```{r}
#Model 1
#random intercept for participant - can have different baseline levels of theta
#random slope for participant - can have different slopes between 6mo and 2yr
#with absolute power other regions
lmm_rand_time_otherROIs <- lmer(Theta_absoluteT ~ otherROIs * timepoint + (1 + timepoint| sid), data = dat_otherRegions)

summary(lmm_rand_time_otherROIs)
confint(lmm_rand_time_otherROIs)
AIC(lmm_rand_time_otherROIs)
BIC(lmm_rand_time_otherROIs)
logLik(lmm_rand_time_otherROIs)


#Model 2
#no interaction terms with num segments EEG included as fixed effect - with relative power
options(scipen = 999)
lmm_rand_time_adv_otherROIs <- lmer(Theta_absoluteT ~ sex + maternal_yrsofed + income_per_day_impute + HAZ_6mo_impute + PSS_sum_score  + num_segs_EEG + otherROIs * timepoint + (1 + timepoint| sid), data = dat_otherRegions)
summary(lmm_rand_time_adv_otherROIs)
confint(lmm_rand_time_adv_otherROIs)
AIC(lmm_rand_time_adv_otherROIs)
BIC(lmm_rand_time_adv_otherROIs)
logLik(lmm_rand_time_adv_otherROIs)



#Model 3
#interaction terms with number EEG segs as fixed effect - with relative power
lmm_rand_time_adv_inter_otherROIs <- lmer(Theta_absoluteT ~ num_segs_EEG + sex*timepoint + maternal_yrsofed*timepoint + income_per_day_impute*timepoint + HAZ_6mo_impute*timepoint + PSS_sum_score*timepoint + otherROIs*timepoint + (1 + timepoint| sid), data = dat_otherRegions)
summary(lmm_rand_time_adv_inter_otherROIs)
confint(lmm_rand_time_adv_inter_otherROIs)
nobs(lmm_rand_time_adv_inter_otherROIs)

AIC(lmm_rand_time_adv_inter_otherROIs)
BIC(lmm_rand_time_adv_inter_otherROIs)
logLik(lmm_rand_time_adv_inter_otherROIs)




```

#modeling with relative power: other ROIs
```{r}
#Model 1
#random intercept for participant - can have different baseline levels of theta
#random slope for participant - can have different slopes between 6mo and 2yr
#with absolute power other regions
lmm_rand_time_otherROIs_rel <- lmer(Theta_relativeT ~ otherROIs * timepoint + (1 + timepoint| sid), data = dat_otherRegions)

summary(lmm_rand_time_otherROIs_rel)
confint(lmm_rand_time_otherROIs_rel)
AIC(lmm_rand_time_otherROIs_rel)
BIC(lmm_rand_time_otherROIs_rel)
logLik(lmm_rand_time_otherROIs_rel)


#Model 2
#no interaction terms with num segments EEG included as fixed effect - with relative power
options(scipen = 999)
lmm_rand_time_adv_otherROIs_rel <- lmer(Theta_relativeT ~ sex + maternal_yrsofed + income_per_day_impute + HAZ_6mo_impute + PSS_sum_score  + num_segs_EEG + otherROIs * timepoint + (1 + timepoint| sid), data = dat_otherRegions)
summary(lmm_rand_time_adv_otherROIs_rel)
confint(lmm_rand_time_adv_otherROIs_rel)
AIC(lmm_rand_time_adv_otherROIs_rel)
BIC(lmm_rand_time_adv_otherROIs_rel)
logLik(lmm_rand_time_adv_otherROIs_rel)



#Model 3
#interaction terms with number EEG segs as fixed effect - with relative power
lmm_rand_time_adv_inter_otherROIs_rel <- lmer(Theta_relativeT ~ num_segs_EEG + sex*timepoint + maternal_yrsofed*timepoint + income_per_day_impute*timepoint + HAZ_6mo_impute*timepoint + PSS_sum_score*timepoint + otherROIs*timepoint + (1 + timepoint| sid), data = dat_otherRegions)
summary(lmm_rand_time_adv_inter_otherROIs_rel)
confint(lmm_rand_time_adv_inter_otherROIs_rel)
nobs(lmm_rand_time_adv_inter_otherROIs_rel)

AIC(lmm_rand_time_adv_inter_otherROIs_rel)
BIC(lmm_rand_time_adv_inter_otherROIs_rel)
logLik(lmm_rand_time_adv_inter_otherROIs_rel)




```


#regressions with 2yr relative power (for Aim 3)
```{r}
#convert relative frontal dataset to wide
dat_rel_theta_wide <- dat_rel_theta_long %>%
  pivot_wider(
    id_cols = c(sid), # Columns to keep as-is
    names_from = c(ROI, timepoint),        # Columns to create new column names from
    names_sep = "_",     # Separator for new column names
    names_prefix = "relative_",
    values_from = relative_theta_power     # Values to populate the new columns
  )

#merge this dataset with wide dataset with everything else
dat_with_rel <- dat_merged <- dat %>%
  left_join(dat_rel_theta_wide, by = "sid")

#run regressions with IQ, WM, and inhibition as outcomes
options(digits=10)
theta_WPPSI_rel <- lm(fsiq_5yr ~ relative_CentralZ_2yr + relative_FrontalL_2yr + relative_FrontalR_2yr, data = dat_with_rel)
summary(theta_WPPSI_rel)
nobs(theta_WPPSI_rel)

theta_GNG_rel <- lm(GNG_accuracy_all ~ relative_CentralZ_2yr + relative_FrontalL_2yr + relative_FrontalR_2yr, data = dat_with_rel)
summary(theta_GNG_rel)
nobs(theta_GNG_rel)


theta_PALTEA_rel <- lm(PALTEA ~ relative_CentralZ_2yr + relative_FrontalL_2yr + relative_FrontalR_2yr, data = dat_with_rel)
summary(theta_PALTEA_rel)
nobs(theta_PALTEA_rel)


 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI_rel <- summary(theta_WPPSI_rel)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI_rel
p_values_theta_PALTEA_rel <- summary(theta_PALTEA_rel)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA_rel
p_values_theta_GNG_rel <- summary(theta_GNG_rel)$coef[, "Pr(>|t|)"]
p_values_theta_GNG_rel

all_p_values_theta_rel <- c(p_values_theta_WPPSI_rel, p_values_theta_PALTEA_rel, p_values_theta_GNG_rel)
all_p_values_theta_rel
corrected_p_values_theta_rel <- p.adjust(all_p_values_theta_rel, method = "fdr")
corrected_p_values_theta_rel

```

#regressions with absolute and relative power from other regions 
```{r}
#restrict data_otherRegions dataset to just EEG variables
dat_otherRegions_sub <- subset(dat_otherRegions, select = c(sid, timepoint, otherROIs, Theta_absoluteT, Theta_relativeT))

#convert rto wide
dat_otherRegions_sub_wide <- dat_otherRegions_sub %>%
  pivot_wider(
    id_cols = c(sid), # Columns to keep as-is
    names_from = c(otherROIs, timepoint),        # Columns to create new column names from
    names_sep = "_",     # Separator for new column names
    values_from = c(Theta_absoluteT, Theta_relativeT)     # Values to populate the new columns
  )

#merge this dataset with wide dataset with everything else
dat_with_otherROIs <- dat_merged <- dat %>%
  left_join(dat_otherRegions_sub_wide, by = "sid")


#run regressions with IQ, WM, and inhibition as outcomes: Absolute power
options(digits=10)
theta_WPPSI_otherROIs <- lm(fsiq_5yr ~ Theta_absoluteT_Central_2yr + Theta_absoluteT_Frontal_2yr + Theta_absoluteT_Occipital_2yr + Theta_absoluteT_Parietal_2yr + Theta_absoluteT_Temporal_2yr, data = dat_with_otherROIs)
summary(theta_WPPSI_otherROIs)
nobs(theta_WPPSI_otherROIs)

options(digits=10)
theta_GNG_otherROIs <- lm(GNG_accuracy_all ~ Theta_absoluteT_Central_2yr + Theta_absoluteT_Frontal_2yr + Theta_absoluteT_Occipital_2yr + Theta_absoluteT_Parietal_2yr + Theta_absoluteT_Temporal_2yr, data = dat_with_otherROIs)
summary(theta_GNG_otherROIs)
nobs(theta_GNG_otherROIs)


theta_PALTEA_otherROIs <- lm(PALTEA ~ Theta_absoluteT_Central_2yr + Theta_absoluteT_Frontal_2yr + Theta_absoluteT_Occipital_2yr + Theta_absoluteT_Parietal_2yr + Theta_absoluteT_Temporal_2yr, data = dat_with_otherROIs)
summary(theta_PALTEA_otherROIs)
nobs(theta_PALTEA_otherROIs)

 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI_otherROIs <- summary(theta_WPPSI_otherROIs)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI_otherROIs
p_values_theta_PALTEA_otherROIs <- summary(theta_PALTEA_otherROIs)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA_otherROIs
p_values_theta_GNG_otherROIs <- summary(theta_GNG_otherROIs)$coef[, "Pr(>|t|)"]
p_values_theta_GNG_otherROIs

all_p_values_theta_otherROIs <- c(p_values_theta_WPPSI_otherROIs, p_values_theta_PALTEA_otherROIs, p_values_theta_GNG_otherROIs)
all_p_values_theta_otherROIs
corrected_p_values_theta_otherROIs <- p.adjust(all_p_values_theta_otherROIs, method = "fdr")
corrected_p_values_theta_otherROIs



#run regressions with IQ, WM, and inhibition as outcomes: Relative power
options(digits=10)
theta_WPPSI_otherROIs_rel <- lm(fsiq_5yr ~ Theta_relativeT_Central_2yr + Theta_absoluteT_Frontal_2yr + Theta_absoluteT_Occipital_2yr + Theta_absoluteT_Parietal_2yr + Theta_absoluteT_Temporal_2yr, data = dat_with_otherROIs)
summary(theta_WPPSI_otherROIs_rel)
nobs(theta_WPPSI_otherROIs_rel)

options(digits=10)
theta_GNG_otherROIs_rel <- lm(GNG_accuracy_all ~ Theta_relativeT_Central_2yr + Theta_absoluteT_Frontal_2yr + Theta_absoluteT_Occipital_2yr + Theta_absoluteT_Parietal_2yr + Theta_absoluteT_Temporal_2yr, data = dat_with_otherROIs)
summary(theta_GNG_otherROIs_rel)
nobs(theta_GNG_otherROIs_rel)


theta_PALTEA_otherROIs_rel <- lm(PALTEA ~ Theta_relativeT_Central_2yr + Theta_absoluteT_Frontal_2yr + Theta_absoluteT_Occipital_2yr + Theta_absoluteT_Parietal_2yr + Theta_absoluteT_Temporal_2yr, data = dat_with_otherROIs)
summary(theta_PALTEA_otherROIs_rel)
nobs(theta_PALTEA_otherROIs_rel)

 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI_otherROIs_rel <- summary(theta_WPPSI_otherROIs_rel)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI_otherROIs_rel
p_values_theta_PALTEA_otherROIs_rel <- summary(theta_PALTEA_otherROIs_rel)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA_otherROIs_rel
p_values_theta_GNG_otherROIs_rel <- summary(theta_GNG_otherROIs_rel)$coef[, "Pr(>|t|)"]
p_values_theta_GNG_otherROIs_rel

all_p_values_theta_otherROIs_rel <- c(p_values_theta_WPPSI_otherROIs_rel, p_values_theta_PALTEA_otherROIs_rel, p_values_theta_GNG_otherROIs_rel)
all_p_values_theta_otherROIs_rel
corrected_p_values_theta_otherROIs_rel <- p.adjust(all_p_values_theta_otherROIs_rel, method = "fdr")
corrected_p_values_theta_otherROIs_rel
```

#re-format individual trial-level data to get it in format of columns are participants and rows are trials: 6mo Cz
```{r}
library(readr)
library(dplyr)
library(purrr)
library(writexl)

# Define the parent directory containing participant folders
parent_dir <- "/Users/eilysullivan/BCH Dropbox/Eileen Sullivan/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/6mo_processed_set/generatePower_indivtrials/bandpower_IndivTrials"

# List all participant folders
folders <- list.dirs(parent_dir, recursive = FALSE)

# Initialize an empty list to store data
average_chan_list <- list()
row_column <- NULL  # Placeholder for the "Row" column

# Loop through each folder
for (folder in folders) {
  # Extract participant ID from folder name
  participant_id <- basename(folder)
  
  # Find the relevant file
  files <- list.files(folder, pattern = "_indivBandpower_Theta_Central_Z\\.csv$", full.names = TRUE)
  
  if (length(files) == 1) {
    # Read the file
    data <- read_csv(files[1], show_col_types=FALSE)
    
    # Extract "Row" column if it's not already set
    if (is.null(row_column)) {
      row_column <- data$Row
    }
    
    # Extract "AverageChan" column
    average_chan_list[[participant_id]] <- data$AverageChan
  } else {
    message("Skipping folder: ", folder, " - No or multiple matching files found.")
  }
}

# Determine the maximum column length
max_length <- max(map_int(average_chan_list, length))

# Pad each "AverageChan" column with NA to match the maximum length
average_chan_padded <- map(average_chan_list, ~c(.x, rep(NA, max_length - length(.x))))

# Combine all AverageChan columns into a dataframe
average_chan_df <- bind_cols(average_chan_padded)

# Add the "Row" column as the first column
final_df_Cz_6mo <- tibble(Row = c(row_column, rep(NA, max_length - length(row_column)))) %>%
  bind_cols(average_chan_df)

final_df_Cz_6mo <- final_df_Cz_6mo %>%
  slice(-n())

# Write to a new Excel file
write_xlsx(final_df_Cz_6mo, "Indiv_trial_data/Central_Z_6mo_Indivtrial_Theta.xlsx")

```

#calculate internal consistency: 6mo Cz
```{r}
# Remove the "Row" column from your dataset
final_df_Cz_6mo_clean <- final_df_Cz_6mo %>%
  select(-Row)

# Function to calculate split-half reliability for one participant
split_half_reliability <- function(trials) {
  # Remove NAs
  trials <- na.omit(trials)
  
  # Check if there are enough trials for computation
  if (length(trials) < 2) return(NA)  # Reliability can't be calculated with < 2 trials

  # Split trials into two random halves
  set.seed(456)  # For reproducibility
  split_indices <- sample(seq_along(trials))  # Random shuffle of indices
  mid_point <- ceiling(length(split_indices) / 2)  # Determine split point
  half1 <- trials[split_indices[1:mid_point]]  # First half
  half2 <- trials[split_indices[(mid_point + 1):length(split_indices)]]  # Second half

  # Check if halves are of equal length
  if (length(half1) != length(half2)) {
    # Add NA padding to the shorter half to match lengths
    if (length(half1) > length(half2)) {
      half2 <- c(half2, rep(NA, length(half1) - length(half2)))
    } else {
      half1 <- c(half1, rep(NA, length(half2) - length(half1)))
    }
  }

  # Compute means of each half
  mean_half1 <- mean(half1, na.rm = TRUE)
  mean_half2 <- mean(half2, na.rm = TRUE)

  # Compute correlation and reliability
  corr <- cor(half1, half2, use = "pairwise.complete.obs")
  if (is.na(corr)) return(NA)

  reliability <- 2 * corr / (1 + corr)
  return(reliability)
}

# Apply the function to each participant's data
reliabilities <- sapply(final_df_Cz_6mo_clean, split_half_reliability)

# Compute the mean reliability across participants
mean_reliability <- mean(reliabilities, na.rm = TRUE)
print(paste("Mean Split-Half Reliability:", mean_reliability))

# Optional: Inspect reliabilities for individual participants
print(reliabilities)
hist(reliabilities)

test_trials <- na.omit(final_df_Cz_6mo_clean[[2]])  # Replace with participant column
set.seed(123)
split_indices <- sample(seq_along(test_trials))
mid_point <- ceiling(length(split_indices) / 2)
half1 <- test_trials[split_indices[1:mid_point]]
half2 <- test_trials[split_indices[(mid_point + 1):length(split_indices)]]
print(cor(half1, half2, use = "pairwise.complete.obs"))

# Calculate Cronbach's alpha
alpha_result <- psych::alpha(final_df_Cz_6mo_clean, na.rm = TRUE)

# View the result
print(alpha_result$total)  # Overall Cronbach's alpha

```


#re-format individual trial-level data to get it in format of columns are participants and rows are trials: 6mo FrontalL
```{r}
library(readr)
library(dplyr)
library(purrr)
library(writexl)

# Define the parent directory containing participant folders
parent_dir <- "/Users/eilysullivan/BCH Dropbox/Eileen Sullivan/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/6mo_processed_set/generatePower_indivtrials/bandpower_IndivTrials"

# List all participant folders
folders <- list.dirs(parent_dir, recursive = FALSE)

# Initialize an empty list to store data
average_chan_list <- list()
row_column <- NULL  # Placeholder for the "Row" column

# Loop through each folder
for (folder in folders) {
  # Extract participant ID from folder name
  participant_id <- basename(folder)
  
  # Find the relevant file
  files <- list.files(folder, pattern = "_indivBandpower_Theta_Frontal_L\\.csv$", full.names = TRUE)
  
  if (length(files) == 1) {
    # Read the file
    data <- read_csv(files[1], show_col_types=FALSE)
    
    # Extract "Row" column if it's not already set
    if (is.null(row_column)) {
      row_column <- data$Row
    }
    
    # Extract "AverageChan" column
    average_chan_list[[participant_id]] <- data$AverageChan
  } else {
    message("Skipping folder: ", folder, " - No or multiple matching files found.")
  }
}

# Determine the maximum column length
max_length <- max(map_int(average_chan_list, length))

# Pad each "AverageChan" column with NA to match the maximum length
average_chan_padded <- map(average_chan_list, ~c(.x, rep(NA, max_length - length(.x))))

# Combine all AverageChan columns into a dataframe
average_chan_df <- bind_cols(average_chan_padded)

# Add the "Row" column as the first column
final_df_FL_6mo <- tibble(Row = c(row_column, rep(NA, max_length - length(row_column)))) %>%
  bind_cols(average_chan_df)

final_df_FL_6mo <- final_df_FL_6mo %>%
  slice(-n())

# Write to a new Excel file
write_xlsx(final_df_FL_6mo, "Indiv_trial_data/Frontal_L_6mo_Indivtrial_Theta.xlsx")

```


#re-format individual trial-level data to get it in format of columns are participants and rows are trials: 6mo FrontalR
```{r}
library(readr)
library(dplyr)
library(purrr)
library(writexl)

# Define the parent directory containing participant folders
parent_dir <- "/Users/eilysullivan/BCH Dropbox/Eileen Sullivan/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/6mo_processed_set/generatePower_indivtrials/bandpower_IndivTrials"

# List all participant folders
folders <- list.dirs(parent_dir, recursive = FALSE)

# Initialize an empty list to store data
average_chan_list <- list()
row_column <- NULL  # Placeholder for the "Row" column

# Loop through each folder
for (folder in folders) {
  # Extract participant ID from folder name
  participant_id <- basename(folder)
  
  # Find the relevant file
  files <- list.files(folder, pattern = "_indivBandpower_Theta_Frontal_R\\.csv$", full.names = TRUE)
  
  if (length(files) == 1) {
    # Read the file
    data <- read_csv(files[1], show_col_types=FALSE)
    
    # Extract "Row" column if it's not already set
    if (is.null(row_column)) {
      row_column <- data$Row
    }
    
    # Extract "AverageChan" column
    average_chan_list[[participant_id]] <- data$AverageChan
  } else {
    message("Skipping folder: ", folder, " - No or multiple matching files found.")
  }
}

# Determine the maximum column length
max_length <- max(map_int(average_chan_list, length))

# Pad each "AverageChan" column with NA to match the maximum length
average_chan_padded <- map(average_chan_list, ~c(.x, rep(NA, max_length - length(.x))))

# Combine all AverageChan columns into a dataframe
average_chan_df <- bind_cols(average_chan_padded)

# Add the "Row" column as the first column
final_df_FR_6mo <- tibble(Row = c(row_column, rep(NA, max_length - length(row_column)))) %>%
  bind_cols(average_chan_df)

final_df_FR_6mo <- final_df_FR_6mo %>%
  slice(-n())

# Write to a new Excel file
write_xlsx(final_df_FR_6mo, "Indiv_trial_data/Frontal_R_6mo_Indivtrial_Theta.xlsx")


#calculate internal consistency: 6mo Frontal Left
```

```{r}
#Cronbachs alpha 6mo Cz:
alpha_result_Cz <- psych::alpha(final_df_Cz_6mo_clean, na.rm = TRUE, check.keys = TRUE)

# View the result
print(alpha_result_Cz$total)  # Overall Cronbach's alpha

# Calculate Cronbach's alpha 6mo FL: 
final_df_FL_6mo_clean <- final_df_FL_6mo %>%
  select(-Row)
alpha_result_FL <- psych::alpha(final_df_FL_6mo_clean, na.rm = TRUE, check.keys = TRUE)
# View the result
print(alpha_result_FL$total)  # Overall Cronbach's alpha

#Cronbachs alpha: 6mo FR
final_df_FR_6mo_clean <- final_df_FR_6mo %>%
  select(-Row)
# Calculate Cronbach's alpha 6mo Frontal Left
alpha_result_FR <- psych::alpha(final_df_FR_6mo_clean, na.rm = TRUE, check.keys = TRUE)
# View the result
print(alpha_result_FR$total)  # Overall Cronbach's alpha

```
#2yo 

#re-format individual trial-level data to get it in format of columns are participants and rows are trials: 2yr Cz
```{r}
library(readr)
library(dplyr)
library(purrr)
library(writexl)

# Define the parent directory containing participant folders
parent_dir <- "/Users/eilysullivan/BCH Dropbox/Eileen Sullivan/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/2yr_processed_set/generatePower_individualtrials/bandpower_IndivTrials"

# List all participant folders
folders <- list.dirs(parent_dir, recursive = FALSE)

# Initialize an empty list to store data
average_chan_list <- list()
row_column <- NULL  # Placeholder for the "Row" column

# Loop through each folder
for (folder in folders) {
  # Extract participant ID from folder name
  participant_id <- basename(folder)
  
  # Find the relevant file
  files <- list.files(folder, pattern = "_indivBandpower_Theta_Central_Z\\.csv$", full.names = TRUE)
  
  if (length(files) == 1) {
    # Read the file
    data <- read_csv(files[1], show_col_types=FALSE)
    
    # Extract "Row" column if it's not already set
    if (is.null(row_column)) {
      row_column <- data$Row
    }
    
    # Extract "AverageChan" column
    average_chan_list[[participant_id]] <- data$AverageChan
  } else {
    message("Skipping folder: ", folder, " - No or multiple matching files found.")
  }
}

# Determine the maximum column length
max_length <- max(map_int(average_chan_list, length))

# Pad each "AverageChan" column with NA to match the maximum length
average_chan_padded <- map(average_chan_list, ~c(.x, rep(NA, max_length - length(.x))))

# Combine all AverageChan columns into a dataframe
average_chan_df <- bind_cols(average_chan_padded)

# Add the "Row" column as the first column
final_df_Cz_2yr <- tibble(Row = c(row_column, rep(NA, max_length - length(row_column)))) %>%
  bind_cols(average_chan_df)

final_df_Cz_2yr <- final_df_Cz_2yr %>%
  slice(-n())

# Write to a new Excel file
write_xlsx(final_df_Cz_2yr, "Indiv_trial_data/Central_Z_2yr_Indivtrial_Theta.xlsx")

```




#re-format individual trial-level data to get it in format of columns are participants and rows are trials: 2yr Frontal L
```{r}
library(readr)
library(dplyr)
library(purrr)
library(writexl)

# Define the parent directory containing participant folders
parent_dir <- "/Users/eilysullivan/BCH Dropbox/Eileen Sullivan/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/2yr_processed_set/generatePower_individualtrials/bandpower_IndivTrials"

# List all participant folders
folders <- list.dirs(parent_dir, recursive = FALSE)

# Initialize an empty list to store data
average_chan_list <- list()
row_column <- NULL  # Placeholder for the "Row" column

# Loop through each folder
for (folder in folders) {
  # Extract participant ID from folder name
  participant_id <- basename(folder)
  
  # Find the relevant file
  files <- list.files(folder, pattern = "_indivBandpower_Theta_Frontal_L\\.csv$", full.names = TRUE)
  
  if (length(files) == 1) {
    # Read the file
    data <- read_csv(files[1], show_col_types=FALSE)
    
    # Extract "Row" column if it's not already set
    if (is.null(row_column)) {
      row_column <- data$Row
    }
    
    # Extract "AverageChan" column
    average_chan_list[[participant_id]] <- data$AverageChan
  } else {
    message("Skipping folder: ", folder, " - No or multiple matching files found.")
  }
}

# Determine the maximum column length
max_length <- max(map_int(average_chan_list, length))

# Pad each "AverageChan" column with NA to match the maximum length
average_chan_padded <- map(average_chan_list, ~c(.x, rep(NA, max_length - length(.x))))

# Combine all AverageChan columns into a dataframe
average_chan_df <- bind_cols(average_chan_padded)

# Add the "Row" column as the first column
final_df_FL_2yr <- tibble(Row = c(row_column, rep(NA, max_length - length(row_column)))) %>%
  bind_cols(average_chan_df)

final_df_FL_2yr <- final_df_FL_2yr %>%
  slice(-n())

# Write to a new Excel file
write_xlsx(final_df_FL_2yr, "Indiv_trial_data/Frontal_L_2yr_Indivtrial_Theta.xlsx")

```


#re-format individual trial-level data to get it in format of columns are participants and rows are trials: 2yr Frontal R
```{r}
library(readr)
library(dplyr)
library(purrr)
library(writexl)

# Define the parent directory containing participant folders
parent_dir <- "/Users/eilysullivan/BCH Dropbox/Eileen Sullivan/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/2yr_processed_set/generatePower_individualtrials/bandpower_IndivTrials"

# List all participant folders
folders <- list.dirs(parent_dir, recursive = FALSE)

# Initialize an empty list to store data
average_chan_list <- list()
row_column <- NULL  # Placeholder for the "Row" column

# Loop through each folder
for (folder in folders) {
  # Extract participant ID from folder name
  participant_id <- basename(folder)
  
  # Find the relevant file
  files <- list.files(folder, pattern = "_indivBandpower_Theta_Frontal_R\\.csv$", full.names = TRUE)
  
  if (length(files) == 1) {
    # Read the file
    data <- read_csv(files[1], show_col_types=FALSE)
    
    # Extract "Row" column if it's not already set
    if (is.null(row_column)) {
      row_column <- data$Row
    }
    
    # Extract "AverageChan" column
    average_chan_list[[participant_id]] <- data$AverageChan
  } else {
    message("Skipping folder: ", folder, " - No or multiple matching files found.")
  }
}

# Determine the maximum column length
max_length <- max(map_int(average_chan_list, length))

# Pad each "AverageChan" column with NA to match the maximum length
average_chan_padded <- map(average_chan_list, ~c(.x, rep(NA, max_length - length(.x))))

# Combine all AverageChan columns into a dataframe
average_chan_df <- bind_cols(average_chan_padded)

# Add the "Row" column as the first column
final_df_FR_2yr <- tibble(Row = c(row_column, rep(NA, max_length - length(row_column)))) %>%
  bind_cols(average_chan_df)

final_df_FR_2yr <- final_df_FR_2yr %>%
  slice(-n())

# Write to a new Excel file
write_xlsx(final_df_FR_2yr, "Indiv_trial_data/Frontal_R_2yr_Indivtrial_Theta.xlsx")


```

```{r}
#Cronbachs alpha 2yr Cz:
final_df_Cz_2yr_clean <- final_df_Cz_2yr %>%
  select(-Row)
alpha_result_Cz_2yr <- psych::alpha(final_df_Cz_2yr_clean, na.rm = TRUE, check.keys = TRUE)

# View the result
print(alpha_result_Cz_2yr$total)  # Overall Cronbach's alpha

# Calculate Cronbach's alpha 2yr FL: 
final_df_FL_2yr_clean <- final_df_FL_2yr %>%
  select(-Row)
alpha_result_FL_2yr <- psych::alpha(final_df_FL_2yr_clean, na.rm = TRUE, check.keys = TRUE)
# View the result
print(alpha_result_FL_2yr$total)  # Overall Cronbach's alpha

#Cronbachs alpha: 2yr FR
final_df_FR_2yr_clean <- final_df_FR_2yr %>%
  select(-Row)
# Calculate Cronbach's alpha 2yr Frontal Left
alpha_result_FR_2yr <- psych::alpha(final_df_FR_2yr_clean, na.rm = TRUE, check.keys = TRUE)
# View the result
print(alpha_result_FR_2yr$total)  # Overall Cronbach's alpha

```

