---
title: "Crypto_EEG_CANTAB_paper"
author: "Eily Sullivan"
date: "5/2/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages, include=FALSE}

rm(list=ls())
library(tidyverse)
library(haven)
library(dplyr)
library(foreign)
library(lmtest)
library(lme4)
library(lmerTest)
library(ggplot2)
#install.packages("standardize")
library(standardize)
library(ggthemes)
library(texreg)
library(rockchalk)
library(stargazer)
#install.packages("Hmisc")
library(Hmisc)
library(writexl)
library("readxl")
library(corrplot)
#install.packages("correlation_matrix")
library(psych)
library(tidyr)
#install.packages("broom")  # For tidying the model summary
#library(broom)
#install.packages("officer")  # For exporting to Word
library(officer)
#install.packages("imputeTS")



```


#load the dataset
```{r}

#full datapath: FINAL_Crypto_DATASET_EEG_CANTAB <- read_excel("Dropbox (BCH)/BEAN baseline EEG attention/1_PAPER_RS_EEG_CANTAB/Datasets/FINAL_Crypto_DATASET_EEG_CANTAB.xlsx")

dat <- FINAL_Crypto_DATASET_EEG_CANTAB <- read_excel("FINAL_Crypto_DATASET_EEG_CANTAB.xlsx")

#remove last row
dat <- dat[-nrow(dat), ]

#make categorical variables factors
dat$cohort <- as.factor(dat$cohort)

dat$sex <- as.factor(dat$sex)
summary(dat$sex)

dat$child_use_screens <- as.factor(dat$child_use_screens)

#make ages numeric
dat$age_mo_2yr_EEG <- as.numeric(dat$age_mo_2yr_EEG)


```

#log transform EEG variables
```{r}

dat$Log_central_z_T_6mo <- log10(dat$central_z_T_6mo)
dat$Log_frontal_l_T_6mo <- log10(dat$frontal_l_T_6mo)
dat$Log_frontal_r_T_6mo <- log10(dat$frontal_r_T_6mo)

dat$Log_central_z_B_6mo <- log10(dat$central_z_B_6mo)
dat$Log_frontal_l_B_6mo <- log10(dat$frontal_l_B_6mo)
dat$Log_frontal_r_B_6mo <- log10(dat$frontal_r_B_6mo)

dat$Log_central_z_T_2yr <- log10(dat$central_z_T_2yr)
dat$Log_frontal_l_T_2yr <- log10(dat$frontal_l_T_2yr)
dat$Log_frontal_r_T_2yr <- log10(dat$frontal_r_T_2yr)

dat$Log_central_z_B_2yr <- log10(dat$central_z_B_2yr)
dat$Log_frontal_l_B_2yr <- log10(dat$frontal_l_B_2yr)
dat$Log_frontal_r_B_2yr <- log10(dat$frontal_r_B_2yr)

#LOG descriptives
summary(dat$Log_central_z_T_6mo)
#below 3SD would be 0.53597

summary(dat$Log_frontal_l_T_6mo)
#below 3 SD would 1.1976 - 0.52498 = 0.6726



summary(dat$Log_frontal_r_T_6mo)
#below 3SD would be 0.687


summary(dat$Log_central_z_T_2yr)
#below 3SD would be 0.67737

summary(dat$Log_frontal_l_T_2yr)
#below 3SD would be 0.8533239

summary(dat$Log_frontal_r_T_2yr)
#below 3SD would be 0.8019

#remove observations if below 3SD from mean 
below_6mo_central_z_rows <- dat[dat$Log_central_z_T_6mo < 0.53597, ]
print(below_6mo_central_z_rows)

below_6mo_frontal_l_rows <- dat[dat$Log_frontal_l_T_6mo < 0.6726, ]
print(below_6mo_frontal_l_rows)

below_6mo_frontal_r_rows <- dat[dat$Log_frontal_l_T_6mo < 0.687, ]
print(below_6mo_frontal_r_rows)


#2yo - no one to be removed
below_2yr_central_z_rows <- dat[dat$Log_central_z_T_2yr < 0.67737, ]
print(below_2yr_central_z_rows)

below_2yr_frontal_l_rows <- dat[dat$Log_frontal_l_T_2yr < 0.8533239, ]
print(below_2yr_frontal_l_rows)

below_2yr_frontal_r_rows <- dat[dat$Log_frontal_l_T_2yr < 0.8019, ]
print(below_2yr_frontal_r_rows)


#no outliers for 2yr
write_xlsx(dat, "EEG_theta_data_wide.xlsx")





```

#calculate number of families living in poverty
```{r}
dat$income_to_needs_per_day <- dat$income_to_needs / 30.437
summary(dat$income_to_needs_per_day)

dat$extreme_poverty <- cut(dat$income_to_needs_per_day,
                       breaks=c(-Inf, 167.42, Inf),
                       labels=c('poverty', 'not_poverty'))

summary(dat$extreme_poverty)
108/(108+8)


#calculate income per day 
dat$income_per_day <- dat$income / 30.437
summary(dat$income_per_day)


#impute HAZ
dat$HAZ_6mo_impute <- impute(dat$HAZ_6mo, method = "mean")
summary(dat$HAZ_6mo)

#impute income
dat$income_per_day_impute <- impute(dat$income_per_day, method = "mean")




```

#descriptive statistics environmental variables
```{r}

summary(dat$sex)

summary(dat$age_mo_6mo_EEG)
sd(dat$age_mo_6mo_EEG, na.rm=TRUE)
ggplot(dat, aes(x=age_mo_6mo_EEG)) + 
  geom_histogram() 

summary(dat$age_mo_2yr_EEG)
sd(dat$age_mo_2yr_EEG, na.rm=TRUE)
ggplot(dat, aes(x=age_mo_2yr_EEG)) + 
  geom_histogram() 

summary(dat$age_mo_5yr_cog_ass)
sd(dat$age_mo_5yr_cog_ass, na.rm=TRUE)
ggplot(dat, aes(x=age_mo_5yr_cog_ass)) + 
  geom_histogram() 

summary(dat$FCI_6mo)
sd(dat$FCI_6mo, na.rm=TRUE)
ggplot(dat, aes(x=FCI_6mo)) + 
  geom_histogram() 

summary(dat$paternal_yrsofed)
sd(dat$paternal_yrsofed, na.rm=TRUE)
ggplot(dat, aes(x=paternal_yrsofed)) + 
  geom_histogram() 

summary(dat$maternal_yrsofed)
sd(dat$maternal_yrsofed, na.rm=TRUE)
ggplot(dat, aes(x=maternal_yrsofed)) + 
  geom_histogram() 

summary(dat$income)
sd(dat$income, na.rm=TRUE)
ggplot(dat, aes(x=income)) + 
  geom_histogram() 


summary(dat$income_to_needs)
sd(dat$income_to_needs, na.rm=TRUE)
ggplot(dat, aes(x=income_to_needs)) + 
  geom_histogram() 

summary(dat$income_to_needs_per_day)
sd(dat$income_to_needs_per_day, na.rm=TRUE)
ggplot(dat, aes(x=income_to_needs_per_day)) + 
  geom_histogram() 


summary(dat$income_per_day)
sd(dat$income_per_day)

summary(dat$HAZ_6mo)
sd(dat$HAZ_6mo, na.rm=TRUE)
ggplot(dat, aes(x=HAZ_6mo)) + 
  geom_histogram() 

summary(dat$social_support_6mo)
sd(dat$social_support_6mo, na.rm=TRUE)
ggplot(dat, aes(x=social_support_6mo)) + 
  geom_histogram() 

summary(dat$PSS_sum_score)
sd(dat$PSS_sum_score, na.rm=TRUE)
ggplot(dat, aes(x=PSS_sum_score)) + 
  geom_histogram() 

```


#calculate number of stunted children 
```{r}

dat$stunted <- cut(dat$HAZ_6mo,
                       breaks=c(-Inf, -2, Inf),
                       labels=c('stunted', 'not_stunted'))

summary(dat$stunted)
18/(93+18)

```

#Calculate # of mothers with low, medium, high stress 
```{r}
#0-13 = low stress

#14-26 = moderate stress

#27-40 = moderate stress 

dat$PSS_categorites <- cut(dat$PSS_sum_score,
                       breaks=c(0, 13, 26, 40),
                       labels=c('Low', "Moderate", 'High'))

summary(dat$PSS_categorites)
22/115
78/115
15/115

```


#Theta EEG descriptives
```{r}
summary(dat$central_z_T_6mo)
sd(dat$central_z_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=central_z_T_6mo)) + 
  geom_histogram() 


summary(dat$frontal_l_T_6mo)
sd(dat$frontal_l_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=frontal_l_T_6mo)) + 
  geom_histogram() 

summary(dat$frontal_r_T_6mo)
sd(dat$frontal_r_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=frontal_r_T_6mo)) + 
  geom_histogram() 


summary(dat$central_z_T_2yr)
sd(dat$central_z_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=central_z_T_2yr)) + 
  geom_histogram() 

summary(dat$frontal_l_T_2yr)
sd(dat$frontal_l_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=frontal_l_T_2yr)) + 
  geom_histogram() 

summary(dat$frontal_r_T_2yr)
sd(dat$frontal_r_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=frontal_r_T_2yr)) + 
  geom_histogram() 


#LOG descriptives
summary(dat$Log_central_z_T_6mo)
sd(dat$Log_central_z_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=Log_central_z_T_6mo)) + 
  geom_histogram() 
#below 3SD would be 0.53597

summary(dat$Log_frontal_l_T_6mo)
sd = sd(dat$Log_frontal_l_T_6mo, na.rm=TRUE)
3*sd
ggplot(dat, aes(x=Log_frontal_l_T_6mo)) + 
  geom_histogram() 
#below 3 SD would 1.1976 - 0.52498 = 0.6726



summary(dat$Log_frontal_r_T_6mo)
sd(dat$Log_frontal_r_T_6mo, na.rm=TRUE)
ggplot(dat, aes(x=Log_frontal_r_T_6mo)) + 
  geom_histogram() 
#below 3SD would be 0.687


summary(dat$Log_central_z_T_2yr)
sd(dat$Log_central_z_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=Log_central_z_T_2yr)) + 
  geom_histogram() 
#below 3SD would be 0.67737

summary(dat$Log_frontal_l_T_2yr)
sd(dat$Log_frontal_l_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=Log_frontal_l_T_2yr)) + 
  geom_histogram() 
#below 3SD would be 0.8533239

summary(dat$Log_frontal_r_T_2yr)
sd(dat$Log_frontal_r_T_2yr, na.rm=TRUE)
ggplot(dat, aes(x=Log_frontal_r_T_2yr)) + 
  geom_histogram() 
#below 3SD would be 0.8019



```



#Screen use descriptives
```{r}

summary(dat$child_use_screens)
histogram(dat$child_use_screens)

summary(dat$minutes_on_screens)
sd(dat$minutes_on_screens, na.rm=TRUE)
ggplot(dat, aes(x=minutes_on_screens)) + 
  geom_histogram() 


```

#CANTAB descriptive statistics
```{r}

summary(dat$SWMTE)
ggplot(dat, aes(x=SWMTE)) + 
  geom_histogram() 

summary(dat$PALTEA)
sd(dat$PALTEA, na.rm=TRUE)
ggplot(dat, aes(x=PALTEA)) + 
  geom_histogram() 

summary(dat$SSPFTUE)
ggplot(dat, aes(x=SSPFTUE)) + 
  geom_histogram() 

summary(dat$SOCPSMMT)
ggplot(dat, aes(x=SOCPSMMT)) + 
  geom_histogram() 

```
#GNG descriptive statistics
```{r}
summary(dat$GNG_RT_avg)

ggplot(dat, aes(x=GNG_RT_avg)) + 
  geom_histogram() 

summary(dat$GNG_accuracy_all)
sd(dat$GNG_accuracy_all, na.rm=TRUE)

ggplot(dat, aes(x=GNG_accuracy_all)) + 
  geom_histogram() 


```


#WPPSI descriptive statistics
```{r}
summary(dat$viq_5yr)
ggplot(dat, aes(x=viq_5yr)) + 
  geom_histogram() 

summary(dat$piq_5yr)
ggplot(dat, aes(x=piq_5yr)) + 
  geom_histogram() 

summary(dat$fsiq_5yr)
sd(dat$fsiq_5yr, na.rm=TRUE)
ggplot(dat, aes(x=fsiq_5yr)) + 
  geom_histogram() 

cor.test(dat$fsiq_5yr, dat$piq_5yr) #0.89 correlation


```

#Correlation between screen use and cantab?
```{r}
cor.test(dat$minutes_on_screens, dat$PALTEA)
cor.test(dat$minutes_on_screens, dat$SWMTE)
cor.test(dat$minutes_on_screens, dat$SSPFTUE)
cor.test(dat$minutes_on_screens, dat$SOCPSMMT)

cor.test(dat$minutes_on_screens, dat$GNG_accuracy_go)


##no correlation between minutes on screen and any CANTAB outcome

```

#correlations between maternal stress, depression, and social support
```{r}
cor.test(dat$social_support_6mo, dat$EPDS_6mo)
cor.test(dat$PSS_sum_score, dat$EPDS_6mo)
cor.test(dat$PSS_sum_score, dat$social_support_6mo)


```




#Correlation matrices of non-EEG measures 
```{r}
#correlation matrix with everything possible 
corr_matrix_adv_beh <- select (dat, c(age_mo_6mo_EEG, age_mo_2yr_EEG, age_mo_5yr_cog_ass, maternal_yrsofed, paternal_yrsofed, income_to_needs_per_day, HAZ_6mo, FCI_6mo, PSS_sum_score, social_support_6mo, EPDS_6mo, minutes_on_screens, SWMTE, PALTEA, SSPFTUE, SOCPSMMT, viq_5yr, piq_5yr, fsiq_5yr))

#colnames(corr_matrix_NIRS_2y) <- c("RSTG High TP-Novel", "LSTG High TP-Novel", "Child Adversity", "Maternal Adversity", "Environmental Adversity", "Income-to-Needs", "Maternal Education", "HAZ", "Language (Mullen)")


rcorr(as.matrix(corr_matrix_adv_beh), type = c("pearson"))
pval <- corr.test(corr_matrix_adv_beh, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_adv_beh <- cor(corr_matrix_adv_beh, use="pairwise.complete.obs")

corrplot(corr_adv_beh, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")


#correlation matrix with just the selected variables for final models 

#covariates selected - age, sex, income to needs, maternal education
#adversity measures selected - HAZ, FCI
#CANTAB measure selected - PALTEA
#WPPSI measure selected - FSIQ

#correlation matrix with selected variables
corr_matrix_select_adv_beh <- select (dat, c(maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, GNG_accuracy_all, PALTEA,fsiq_5yr))

rcorr(as.matrix(corr_matrix_select_adv_beh), type = c("pearson"))
pval <- corr.test(corr_matrix_select_adv_beh, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_select_adv_beh <- cor(corr_matrix_select_adv_beh, use="pairwise.complete.obs")

corrplot(corr_select_adv_beh, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")



```




#check whether line noise correlated with power
```{r}

#correlation matrix with selected variables
corr_matrix_theta_linenoise <- select (dat, c(central_z_T_6mo, frontal_l_T_6mo, frontal_r_T_6mo, central_z_T_2yr, frontal_l_T_2yr, frontal_r_T_2yr, central_z_B_6mo, frontal_l_B_6mo, frontal_r_B_6mo, central_z_B_2yr, frontal_l_B_2yr, frontal_r_B_2yr, linenoise_reduction_50hz_6mo, linenoise_reduction_50hz_2yr))

corr_matrix_log_theta_linenoise <- select (dat, c(Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, linenoise_reduction_50hz_6mo, linenoise_reduction_50hz_2yr))
corr_matrix_log_theta_linenoise

rcorr(as.matrix(corr_matrix_log_theta_linenoise), type = c("pearson"))
pval <- corr.test(corr_matrix_log_theta_linenoise, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_theta_linenoise <- cor(corr_matrix_log_theta_linenoise, use="pairwise.complete.obs")

corrplot(corr_theta_linenoise, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")

```


#corr matrix with all included variables
```{r}
#correlation matrix with selected variables
corr_matrix_EEG_behavior <- select (dat, c(central_z_T_6mo, frontal_l_T_6mo, frontal_r_T_6mo, central_z_T_2yr, frontal_l_T_2yr, frontal_r_T_2yr, dat$central_z_T_diff, dat$frontal_l_T_diff, dat$frontal_r_T_diff, maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, PALTEA, GNG_accuracy_all, fsiq_5yr))

new_column_names <- c("6mo Frontocentral Theta", "6mo Frontal Left Theta", "6mo Frontal Right Theta", "2yr Frontocentral Theta", "2yr Frontal Left Theta", "2yr Frontal Right Theta", "Maternal Education (yrs)", "Adjusted Family Income", "HAZ", "Maternal Stress", "Working Memory", "Inhibition", "Full-Scale IQ")

colnames(corr_matrix_EEG_behavior) <- new_column_names

rcorr(as.matrix(corr_matrix_EEG_behavior), type = c("pearson"))
pval <- corr.test(corr_matrix_EEG_behavior, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_EEG_behavior <- cor(corr_matrix_EEG_behavior, use="pairwise.complete.obs")

corrplot(corr_EEG_behavior, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")

#try to make it so stars and numbers appear at the same time
#trace(corrplot, edit=TRUE)



corrplot(corr_EEG_behavior, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', type="lower", diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.8, number.cex=0.5, addCoef.col ='black',
         insig = 'label_sig', pch.col = 'black', order = 'original', tl.col="black")






```

#corr matrix with change in theta power 
```{r}
#change in theta power variables
dat$Log_central_z_T_change <- dat$Log_central_z_T_2yr - dat$Log_central_z_T_6mo

dat$Log_frontal_l_T_change <- dat$Log_frontal_l_T_2yr - dat$Log_frontal_l_T_6mo

dat$Log_frontal_r_T_change <- dat$Log_frontal_r_T_2yr - dat$Log_frontal_r_T_6mo



corr_matrix_log_theta_change <- select (dat, c(Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, Log_central_z_T_change, Log_frontal_l_T_change, Log_frontal_r_T_change, maternal_yrsofed, income_to_needs_per_day, HAZ_6mo, PSS_sum_score, PALTEA, GNG_accuracy_all, fsiq_5yr))

rcorr(as.matrix(corr_matrix_log_theta_change), type = c("pearson"))
pval <- corr.test(corr_matrix_log_theta_change, adjust="none")$p
#pval_2y <- corr.test(corr_matrix_NIRS_2y, adjust="fdr")$p
corr_EEG_behavior <- cor(corr_matrix_log_theta_change, use="pairwise.complete.obs")

corrplot(corr_matrix_log_theta_change, tl.cex=0.8, tl.srt=45, p.mat = pval, method = 'color', diag = FALSE, 
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.6,
         insig = 'label_sig', pch.col = 'grey20', order = 'original', tl.col="black")




```

#Aim 1 - LMMs for just theta power 

#create long dataset (convert from wide)
```{r}

#create newdataset with only theta power and rename columns 

dat_EEG_sub <- select (dat, c(sid, Log_central_z_T_6mo, Log_frontal_l_T_6mo, Log_frontal_r_T_6mo, Log_central_z_T_2yr, Log_frontal_l_T_2yr, Log_frontal_r_T_2yr, sex, maternal_yrsofed, income_per_day_impute, HAZ_6mo_impute, PSS_sum_score))

#rename columns 
colnames(dat_EEG_sub) <- c("sid", "6mo_CentralZ", "6mo_FrontalL", "6mo_FrontalR", "2yr_CentralZ", "2yr_FrontalL", "2yr_FrontalR", "sex", "maternal_yrsofed", "income_per_day_impute", "HAZ_6mo_impute", "PSS_sum_score")

library(tidyr)

dat_EEG_long <- pivot_longer(dat_EEG_sub, 
                          cols = -c(sid, sex, maternal_yrsofed, income_per_day_impute, HAZ_6mo_impute, PSS_sum_score), 
                          names_to = c("timepoint", "ROI"), 
                          names_sep = "_", 
                          values_to = "theta_power")

#make 6mo the reference category
dat_EEG_long$timepoint <- factor(dat_EEG_long$timepoint)
summary(dat_EEG_long$timepoint)
dat_EEG_long$timepoint <- relevel(dat_EEG_long$timepoint, ref = "6mo")


#now need to create three separate datasets for each ROI separately 
dat_EEG_long_central_z <- filter(dat_EEG_long, ROI == "CentralZ")
dat_EEG_long_frontal_l <- filter(dat_EEG_long, ROI == "FrontalL")
dat_EEG_long_frontal_r <- filter(dat_EEG_long, ROI == "FrontalR")

```


#Model 1: one big model with just EEG data
```{r}
#random intercept for participant - can have different baseline levels of theta
#random slope for participant - can have different slopes between 6mo and 2yr
lmm_rand_time <- lmer(theta_power ~ ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long)

summary(lmm_rand_time)
confint(lmm_rand_time)
AIC(lmm_rand_time)
BIC(lmm_rand_time)
logLik(lmm_rand_time)


#variance explained by individual variation: 
0.024093 / (0.024093 + 0.037238 + 0.004808)
#variance explained by timepoint:
0.037238 / (0.024093 + 0.037238 + 0.004808)
#variance explained by residual:
0.004808 / (0.024093 + 0.037238 + 0.004808)





```


#Aim 2: add adversity measures to one big model
```{r}

#impute HAZ and income (above)

#no interaction terms
lmm_rand_time_adv <- lmer(theta_power ~ sex + maternal_yrsofed + income_per_day_impute + HAZ_6mo_impute + PSS_sum_score + ROI * timepoint + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time_adv)
confint(lmm_rand_time_adv)
AIC(lmm_rand_time_adv)
BIC(lmm_rand_time_adv)
logLik(lmm_rand_time_adv)



#interaction terms
lmm_rand_time_adv_inter <- lmer(theta_power ~ sex*timepoint + maternal_yrsofed*timepoint + income_per_day_impute*timepoint + HAZ_6mo_impute*timepoint + PSS_sum_score*timepoint + ROI*timepoint + (1 + timepoint| sid), data = dat_EEG_long)
summary(lmm_rand_time_adv_inter)
confint(lmm_rand_time_adv_inter)
nobs(lmm_rand_time_adv_inter)

AIC(lmm_rand_time_adv_inter)
BIC(lmm_rand_time_adv_inter)
logLik(lmm_rand_time_adv_inter)


#worse AIC and BIC when include adversity measures - does not improve the model


```


#Likelihood ratio tests 
```{r}

anova(lmm_rand_time, lmm_rand_time_adv)

anova(lmm_rand_time_adv, lmm_rand_time_adv_inter)

anova(lmm_rand_time, lmm_rand_time_adv_inter)



```


#t-tests for ROI differences within age (probably don't need this)
```{r}

#CREATE AVERAGE OF THETA POWER 
dat$Log_T_6mo_avg <- (dat$Log_central_z_T_6mo + dat$Log_frontal_l_T_6mo + dat$Log_frontal_r_T_6mo)/3
dat$Log_T_2yr_avg <- (dat$Log_central_z_T_2yr + dat$Log_frontal_l_T_2yr + dat$Log_frontal_r_T_2yr)/3

t.test(dat$Log_T_6mo_avg, dat$Log_T_2yr_avg)



#do follow-up t-tests to see whether differences between ROIs within each timepoint

t.test(dat$Log_central_z_T_6mo, dat$Log_frontal_l_T_6mo)
t.test(dat$Log_central_z_T_6mo, dat$Log_frontal_r_T_6mo)
t.test(dat$Log_frontal_l_T_6mo, dat$Log_frontal_r_T_6mo)
#difference between centralZ and two frontal ones (higher in frontal), but no difference between frontal l and r... more lateralized (but not to one side or the other)

t.test(dat$Log_central_z_T_2yr, dat$Log_frontal_l_T_2yr)
t.test(dat$Log_central_z_T_2yr, dat$Log_frontal_r_T_2yr)
t.test(dat$Log_frontal_l_T_2yr, dat$Log_frontal_r_T_2yr)
#difference between frontal l and central z (higher in frontal l) at least before FDR. no difference between central z and frontal r or frontal l or r

#conclusion - maybe more evenly distributed at 2yrs 


#need to correct for multiple comparisons if go this route (correct for 3 or six tests?)

#correcting just within 6mo
p_values_6mo <- c(0.01218509, 0.0141507, 0.910078)
fdr_corrected_p_values_6mo <- p.adjust(p_values_6mo, method = "fdr")
fdr_corrected_p_values_6mo


#correcting just within 2yr
p_values_2yr <- c(0.03264265, 0.189457, 0.3962806)
fdr_corrected_p_values_2yr <- p.adjust(p_values_2yr, method = "fdr")
fdr_corrected_p_values_2yr


#correcting within everything
p_values_all <- c(0.01218509, 0.0141507, 0.910078, 0.03264265, 0.189457, 0.3962806)
fdr_corrected_p_values_all <- p.adjust(p_values_all, method = "fdr")
fdr_corrected_p_values_all



```

#create a bar plot to display these results
```{r}
#relevel ROI
dat_EEG_long$ROI_relev <- factor(dat_EEG_long$ROI, levels = c("FrontalL", "CentralZ", "FrontalR"))

#calculate standard error of each group
summary_data <- dat_EEG_long %>%
  group_by(timepoint, ROI_relev) %>%
  summarise(mean_theta_power = mean(theta_power, na.rm=TRUE),
            se_theta_power = sd(theta_power, na.rm=TRUE) / sqrt(n()))

ggplot(summary_data, aes(x = timepoint, y = mean_theta_power, fill = ROI_relev)) +
  geom_bar(stat = "identity", position = "dodge") + 
geom_errorbar(aes(ymin = mean_theta_power - se_theta_power, ymax = mean_theta_power + se_theta_power),
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(x = "Timepoint", y = "Log Absolute Theta Power") +
  scale_fill_manual(values = c("FrontalL" = "orange", "CentralZ" = "purple", "FrontalR" = "green3"), 
                    labels = c("Frontal Left", "Frontocentral", "Frontal Right")) +
  coord_cartesian(ylim = c(1, 1.4)) +
  theme_minimal() + 
  theme(panel.grid = element_blank()) + 
  labs(fill = "ROI")



ggplot(dat_EEG_long, aes(x = timepoint, y = theta_power, fill = ROI_relev)) +
  geom_bar(stat = "summary", position = "dodge") +
  labs(x = "Timepoint", y = "Log Absolute Theta Power") +
  scale_fill_manual(values = c("FrontalL" = "orange", "CentralZ" = "purple", "FrontalR" = "green3"), 
                    labels = c("Frontal Left", "Frontocentral", "Frontal Right")) +
  theme_minimal() + 
  theme(panel.grid = element_blank()) + 
  labs(fill = "ROI") + 
      coord_cartesian(ylim = c(1, 1.75)) # Adjust y-axis limits without affecting the bars 


  


```



#Aim 3 - regressions predicting cognition

#Regressions: adversity => behavior
```{r}

adv_WPPSI <- lm (fsiq_5yr ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score, data = dat)
summary(adv_WPPSI)
 nobs(adv_WPPSI)
 
 adv_PALTEA <- lm (PALTEA ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score, data = dat)
summary(adv_PALTEA)
 nobs(adv_PALTEA)


adv_GNG <- lm (GNG_accuracy_all ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score + fsiq_5yr, data = dat)
summary(adv_GNG)
nobs(adv_GNG)


adv_PALTEA_control_IQ <- lm (PALTEA ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score + fsiq_5yr, data = dat)
summary(adv_PALTEA_control_IQ)
 nobs(adv_PALTEA)


adv_GNG_control_IQ <- lm (GNG_accuracy_all ~ sex + maternal_yrsofed + income_to_needs_per_day + HAZ_6mo + PSS_sum_score + fsiq_5yr, data = dat)
summary(adv_GNG_control_IQ)
nobs(adv_GNG)


 #correct for multiple comparisons (FDR)
 p_values_adv_WPPSI <- summary(adv_WPPSI)$coef[, "Pr(>|t|)"]
 p_values_adv_WPPSI
p_values_adv_PALTEA <- summary(adv_PALTEA)$coef[, "Pr(>|t|)"]
p_values_adv_PALTEA
p_values_adv_GNG <- summary(adv_GNG)$coef[, "Pr(>|t|)"]

all_p_values_adv <- c(p_values_adv_WPPSI, p_values_adv_PALTEA, p_values_adv_GNG)
all_p_values_adv
corrected_p_values_adv <- p.adjust(all_p_values_adv, method = "fdr")
corrected_p_values_adv


```

#Regressions: 2yr EEG => Cognition

```{r}
options(digits=10)
theta_WPPSI <- lm(fsiq_5yr ~ Log_central_z_T_2yr + Log_frontal_l_T_2yr + Log_frontal_r_T_2yr, data = dat)
summary(theta_WPPSI)
nobs(theta_WPPSI)

theta_GNG <- lm(GNG_accuracy_all ~ Log_central_z_T_2yr + Log_frontal_l_T_2yr + Log_frontal_r_T_2yr, data = dat)
summary(theta_GNG)
nobs(theta_GNG)


theta_PALTEA <- lm(PALTEA ~ Log_central_z_T_2yr + Log_frontal_l_T_2yr + Log_frontal_r_T_2yr, data = dat)
summary(theta_PALTEA)
nobs(theta_PALTEA)


 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI <- summary(theta_WPPSI)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI
p_values_theta_PALTEA <- summary(theta_PALTEA)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA
p_values_theta_GNG <- summary(theta_GNG)$coef[, "Pr(>|t|)"]
p_values_theta_GNG

all_p_values_theta <- c(p_values_theta_WPPSI, p_values_theta_PALTEA, p_values_theta_GNG)
all_p_values_theta
corrected_p_values_theta <- p.adjust(all_p_values_theta, method = "fdr")
corrected_p_values_theta

```


#Regressions: 6mo EEG => Cognition

```{r}
options(digits=10)
theta_WPPSI <- lm(fsiq_5yr ~ Log_central_z_T_6mo + Log_frontal_l_T_6mo + Log_frontal_r_T_6mo, data = dat)
summary(theta_WPPSI)
nobs(theta_WPPSI)

theta_GNG <- lm(GNG_accuracy_all ~ Log_central_z_T_6mo + Log_frontal_l_T_6mo + Log_frontal_r_T_6mo, data = dat)
summary(theta_GNG)
nobs(theta_GNG)


theta_PALTEA <- lm(PALTEA ~ Log_central_z_T_6mo + Log_frontal_l_T_6mo + Log_frontal_r_T_6mo, data = dat)
summary(theta_PALTEA)
nobs(theta_PALTEA)


 #correct for multiple comparisons (FDR)
 p_values_theta_WPPSI <- summary(theta_WPPSI)$coef[, "Pr(>|t|)"]
 p_values_theta_WPPSI
p_values_theta_PALTEA <- summary(theta_PALTEA)$coef[, "Pr(>|t|)"]
p_values_theta_PALTEA
p_values_theta_GNG <- summary(theta_GNG)$coef[, "Pr(>|t|)"]
p_values_theta_GNG

all_p_values_theta <- c(p_values_theta_WPPSI, p_values_theta_PALTEA, p_values_theta_GNG)
all_p_values_theta
corrected_p_values_theta <- p.adjust(all_p_values_theta, method = "fdr")
corrected_p_values_theta

```


#mediation analyses
```{r}
#HAZ => Central_Z theta => FSIQ

#open and run process.R file
#IV to mediator = a path
#mediator to DV = b path (Central Z to IQ)
#direct path IV to DV = c' path (HAZ => IQ)
#total effect = c path


#with just central_z
process (data = dat, y = "fsiq_5yr", x = "HAZ_6mo", m ="central_z_T_2yr", model = 4)
process (data = dat, y = "fsiq_5yr", x = "HAZ_6mo", m ="central_z_T_2yr", model = 4, effsize =1, total =1, stand =1, contrast =1, boot = 10000 , modelbt = 1, seed = 654321)


#with all three ROIs
process(data = dat, y = "fsiq_5yr", x = "HAZ_6mo", m =c("central_z_T_2yr", "frontal_l_T_2yr", "frontal_r_T_2yr"), model = 4, effsize =1, total =1, stand =1, contrast =1, boot = 10000 , modelbt = 1, seed = 654321)


```

